Magnesium modulates phospholipid metabolism to promote bacterial phenotypic resistance to antibiotics

Version 1

Abstract

Abstract

Non-inheritable antibiotic resistance or phenotypic resistance ensures bacterial survival upon antibiotic treatment. However, exogenous factors in promoting phenotypic resistance is poorly defined. Here, we demonstrate that Vibrio alginolyticus are recalcitrant to killing by a broad spectrum of antibiotics under high magnesium. Functional metabolomics demonstrate that magnetism modulates the biosynthesis of fatty acids in increasing the biosynthesis of saturated fatty acids while decreasing unsaturated fatty acids. Exogenous supplementation of fatty acids confirm the role of fatty acids in antibiotic resistance. Furthermore, functional lipidomics reveal that glycerophospholipid metabolism is the major metabolic pathway remodeled by magnetism, where the biosynthesis of PE is decreased but PG is increased. Thus, the membrane composition is altered, leading to increased membrane polarization, and decreased permeability and fluidity. These together reduce the uptake of antibiotics by the bacteria. Thus, our study suggest a previously unrecognized metabolic mechanism by which bacteria escape antibiotic killing by utilizing environmental factor.

Introduction

Non-inheritable antibiotic resistance or phenotypic resistance represent a serious therapeutic challenge for bacterial infectious diseases. Phenotypic resistance does not involve genetic mutations, instead, this type of resistance are transient. Thus, bacteria resume to normal growth after removing the antibiotics. Two extensively studied phenotypic resistance includes biofilm and bacterial persisters (Corona & Martinez, 2013). Biofilms have complex structures that impede the diffusion of antibiotics and contain elements sequestrating antibiotics to prevent their further actions (Ciofu et al, 2022). Moreover, biofilm-forming bacteria and persisters themselves have distinct metabolic state that also significantly negatively influence their susceptibility to antibiotics (Yan & Bassler, 2019). These two types of phenotypic resistance have common phenomena that their growth are retarded or even stopped in the presence of antibiotics (Corona & Martinez, 2013). However, factors promoting bacterial phenotypic resistance that is still proliferate in the presence of antibiotics are poorly defined.

There is a growing body of evidence that metal ions have diverse impact on chemical, physical and physiological phenomena that play roles in development of antibiotic resistance (Booth et al, 2011; Lu et al, 2020; Poole, 2017). One mechanism could involve genetic elements that harbor genes conferring resistance to metals and antibiotics (Poole, 2017) . On the other hand, recent reports suggest that metal cations directly hinder (or enhance) the activity of specific antibiotic drugs (Zhang et al, 2014) and that the metabolic environment impacts bacterial sensitivity to antibiotics (Jiang et al, 2023; Lee & Collins, 2012; Peng et al, 2015; Zhang et al, 2020; Zhao et al, 2021). Light metal ions, such as magnesium, sodium, and potassium, serve as cofactors to many enzymes (Du et al, 2016), some of which could influence drug efficacy. Heavy metal ions, including Cu2+ and Zn2+, confer resistance to antibiotics (Yazdankhah et al, 2014; Zhang et al, 2018). Recent reports suggest that sodium negatively regulates redox states to promote antibiotic resistance in Vibrio alginolyticus (Yang et al, 2018), while actively growing Bacillus subtilis cope with ribosome-targeting antibiotics based on ion flux modulation (Lee et al, 2019). But whether such ions can promote phenotypic resistance is less explored.

Results

Mg2+ promotes resistance to antibiotics

Magnesium is rich in marine and marine aquaculture, where antibiotics are commonly used. We directly investigate the effect of Mg2+ on balofloxacin (BLFX) susceptibility in two strains of Vibrios, V. alginolyticus ATCC33787 and V. parahaemolyticus VP01, which were isolated from marine aquaculture. In the initial experiment, the ionic environment of marine water was simulated by adding 39 mM NaCl, 35 mM Mg2SO4, 7 mM KCl, and 7 mM CaCl2 to LB medium (Table EV1), to create a medium which we designated “artificial seawater” (ASWT). The minimum inhibitory concentrations (MIC) for BLFX in ATCC33787 and VP01 were 54 μg/mL and 25 μg/mL in ASWT and 1 μg/mL and 3 μg/mL in LB medium, respectively (Fig 1A). The role of exogenous NaCl in antibiotic resistance was investigated in a previous study (Yang et al., 2018); and the MIC for BLFX was the same in LB medium or M9 medium plus 7 mM KCl or 7 mM CaCl2 (Fig EV1). However, the MIC for BLFX was higher in M9 medium plus Mg2SO4 or MgCl2 (Fig 1B) and the MIC for BLFX increased with increasing concentration of MgCl2. Specifically, addition of 50 mM or 200 mM MgCl2 increased the MIC for BLFX 200- or 1600-fold, respectively (Fig 1B). Using the Oxford cup method, the zone of inhibition increased with increasing MgCl2 from 15.6 mM – 62.4 mM (Fig 1D). Exogenous MgCl2 also increased MICs for other quinolones (e.g. nalidixic acid, levofloxacin, ciprofloxacin, ofloxacin, and moxifloxacin (Fig 1E) and non-quinolone antibiotics including antibacterial peptides (colistin), macrolides (roxithromycin), tetracyclines (oxytetraycline), β-lactams (ceftriaxone, ceftazidime), and aminoglycosides (amikacine, kanamycin, and gentamicin) (Fig 1F). Such non- specific resistance promoting effects indicate that Mg2+ promotes phenotypic resistance.

Magnesium promotes bacterial resistance to antibiotics.

A. MIC of ATCC 33787 and V. parahaemolyticus to BLFX in SWT or LB medium by microtitre-dilution- method. B. MIC of ATCC 33787 to BLFX in SWT with the replacement of MgCl2 with MgSO4 MgCl2 by microtitre-dilution-method. C. MIC of ATCC 33787 to BLFX in SWT with the indicated concentrations of MgCl2 by microtitre-dilution-method. D. MIC of ATCC 33787 to BLFX in the indicated concentrations of BLFX and MgCl2 by Oxford cup test. E. MIC of ATCC 33787 to other quinolones in SWT with the indicated concentrations of MgCl2 by microtitre-dilution-method. F. MIC of ATCC 33787 to other classes of antibiotics in SWT with the indicated concentrations of MgCl2 by microtitre-dilution-method.

Extracellular Mg2+, intracellular Mg2+ and chelation of antibiotic

Previous reports suggest that chelation of antibiotics by magnesium ions inhibits antibiotic uptake (Deitchman et al, 2018; Lunestad & Goksøyr, 1990). To explore this possibility, the zone of inhibition for balofloxacin was measured with various concentrations of MgCl2, ranging from 0 mM to 200 mM. Marine water usually contains 50 mM MgCl2. Lower concentrations of MgCl2, (0.78, 3.125 or 12.5 mM) did not alter the zone of inhibition, while higher concentration including 50 and 200 mM MgCl2 decreased the zone of inhibition (Fig 2A). Correlation of the MgCl2 concentration-dependent resistance suggests that chelation influences bacterial survival by 25.5% and 27.7% in 50 and 200 mM MgCl2, respectively, but not at lower MgCl2 concentrations (Fig 2B & 2C). Furthermore, intracellular balofloxacin decreased with increasing concentration of MgCl2 (Fig 2D), while exogenous Mg2+ increased intracellular Mg2+ level in a dose-dependent manner, where 50 and 200 mM MgCl2 increased 1.21 and 1.31 mmol/L, respectively (Fig 2E). We also examined rate of efflux and expression of LPS, which contribute to antibiotic resistance (Kobylka et al, 2020; Song et al, 2020). The results show that expression of TolC/tolC and LPS was unaffected by Mg2+ (Fig 2F and 2G). Loss of waaF, lpxA, and lpxC, three key genes involved in LPS biosynthesis, did not influence sensitivity/resistance to balofloxacin (Fig 2H). These data demonstrate that influx of Mg2+ plays roles in balofloxacin resistance.

Mg2+ elevates intracellular Mg2+ and promotes balofloxacin uptake.

A. Oxford cup test of 0, 12.5, 25, 50, 100, 200 μg/mL BLFX to ATCC33787 following by separately incubation with the indicated concentrations of MgCl2 for 5 h. B. Inhibitory curve of data (A). C. Binding efficacy of (A). D. Intracellular BLFX of ATCC 33787 in SWT with the indicated concentrations of MgCl2 and 60 μg/mL BLFX. E. Intracellular Mg2+ of ATCC 33787 in SWT with the indicated concentrations of MgCl2. F. Western blot for abundance of TolC in the presence of MgCl2. G. LPS quantification at indicated concentrations of MgCl2. H. MIC of ATCC 33787 and its mutants ΔwaaF ΔlpxA ΔlpxC in SWT with the indicated concentrations of MgCl2, which is measured by microtitre-dilution-method.

MgCl2 affects bacterial metabolism

Although it is well established that Mg2+ regulates the activity of many metabolic enzymes and strongly influences bacterial metabolism(Garfinkel & Garfinkel, 1985), the mechanism of Mg2+ in promoting antibiotic resistance is less explored. Thus, V. alginolyticus was cultured in M9 medium in the presence of various amounts of MgCl2 (0 mM, 0.78 mM, 3.125 mM, 12.5 mM, 50 mM, or 200 mM), analyzed by GC-MS, and identified 54 metabolites. Five biological and two technical replicates were included for each treatment (Fig EV2). Rank-sum test and a permutation test were used to identify differentially expressed metabolites. Forty-one metabolites were identified (p < 0.05) (Fig 3A), which were presented as a heatmap and Z-value in Fig EV3. Orthogonal partial least square discriminant analysis (OPLS-DA) was conducted that separated the six treatments into three groups: Group 1 included 0 mM, 0.78 mM, and 3.125 mM MgCl2; Group 2 was a singlet of 12.5 mM; Group 3 included 50 mM and 200 mM MgCl2. Component t[1] distinguishes Group 1 from Groups 2 and 3; component t[2] distinguishes Group 3 from Groups 1 and 2 (Fig 3B). Discriminating variables are shown as S-plot, where cut-off values are ≥ 0.05 absolute value of covariance p and 0.5 for correlation p(corr). Six biomarkers/metabolites were selected from component t[1] and p[1]. More specifically, the abundance of cadaverine, urea, palmitic acid, aminoethanol, and fumaric acid were increased, but pyroglutamic acid and glutamic acid were decreased (Fig 3C; Fig EV4). Pathway enrichment analysis suggests that twelve pathways are involved. Notably, one of the pathways is biosynthesis of unsaturated fatty acids (terminology in the software. Actually, it is biosynthesis of fatty acids) (Fig 3D). Interestingly, the abundance of palmitic acid and stearic acid were increased in an Mg2+ dose-dependent manner (Fig 3E; Fig EV5). Moreover, palmitic acid was a crucial biomarker (Fig 3C). Exogenous palmitic acid increased bacterial resistance to balofloxacin (Fig 3F). These results suggest that fatty acid metabolism may be critical to the resistance.

Mg2+-induced metabolomics.

A. Differential metabolomes in the absence or presence of the indicated concentrations of MgCl2. Yellow color and blue color indicate increase and decrease of metabolites relative to the median metabolite level, respectively (see color scale). B. PCA analysis of different concentrations of MgCl2-induced metabolomes. Each dot represents the technical replicate of samples in the plot. C. S-plot generated from OPLS-DA. Predictive component p [1] and correlation p(corr) [1] differentiate 0, 0.78, 3.125 mM MgCl2 from 12.5, 50, 200 mM MgCl2. Predictive component p[2] and correlation p(corr)[2] separate 0, 0.78, 50, 200 mM MgCl2 from 3.125, 12.5 mM MgCl2. The triangle represents metabolites in which candidate biomarkers are marked. D. Enriched pathways by differential abundances of metabolites. E. Scatter plot of palmitic acid and stearic acid in the indicated concentrations of MgCl2, which comes from data (A). F. Synergy analysis for alofloxacin with palmitic acid for V. algilyticus. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (white) and non-synergistic (i.e., additive) agents (purple).

Mg2+ regulates fatty acid biosynthesis

Fatty acids include saturated and unsaturated fatty acids whose biosynthesis shared common biosynthetic pathways from acetyl-CoA to enoyl-ACP, sequentially mediated by fabD or fabH, fabB/F, fabG and fabA/Z. Enoyl-ACP is the direct source to produce unsaturated fatty acids or is metabolized to Acyl-ACP via fabV to produce saturated fatty acids. Unsaturated fatty acids produce saturated fatty acids via tesA/B and yciF. qRT-PCR was used to quantify the expression of involved genes in bacteria treated with different concentration of MgCl2. The expression of 18 genes was increased when treated with 50 or 200 mM MgCl2 (Fig 4A). The increased expression of FabF was confirmed by Western blot (Fig 4B). Expression of tesA, tesB, and yciA, which convert unsaturated to saturated fatty acids, was increased (Fig 4C). The expression of fabA that promotes enoyl-ACP and biosynthesis of unsaturated fatty acids was decreased (Fig 4D), which was confirmed at protein level with Western blot (Fig 4B). Oxazole-2-amine and triclosan inhibit synthesis of fatty acids and decrease bacterial survival in the presence of balofloxacin (Fig 4E). These results indicate that Mg2+ inhibits synthesis of unsaturated fatty acids and promotes synthesis of saturated fatty acids.

Mg2+ promotes biosynthesis of fatty acids.

A. qRT-PCR for expression of fatty acid biosynthesis genes in the absence or presence of MgCl2. B. Western blot for abundance of proteins responsible for conversion to unsatisfied fatty acid biosynthesis or satisfied fatty acid in the absence or presence of MgCl2. C. qRT-PCR for expression of genes encoding conversion from unsatisfied fatty acid biosynthesis to satisfied fatty acid in the absence or presence of MgCl2. D. Diagram for biosynthesis of saturated and unsaturated acids. E. Synergy analysis for balofloxacin with triclosan + oxazole-2- amine for ATCC33787. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (blue) and non-synergistic (i.e., additive) agents (red). F. Diagram for degradation of fatty acids. G. qRT-PCR for expression of genes encoding fatty acid degradation in the absence or presence of the indicated concentrations of MgCl2. H. Western blot for abundance of FadL in the absence or presence of the indicated concentrations of MgCl2. I. qRT-PCR for expression of regulatory genes of fatty acid biosynthesis in the absence or presence of the indicated concentrations of MgCl2. J. Western blot for abundance of FadR in the absence or presence of the indicated concentrations of MgCl2. K and L. Synergy analysis for MgCl2 with BLFX for ΔfadR (L) and ΔarcA (M). Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (while) and non-synergistic (i.e., additive) agents (red). M , N and O qRT-PCR for expression of genes encoding biosynthesis of fatty acids and degradation of fatty acids (M) in ATCC33787 and ΔfadR (N) and ΔarcA (O)in the presence or absence of 200 mM MgCl2.

We also quantified the gene expression of fatty acid degradation (Fig 4F). Interestingly, the expression of genes involved in unsaturated fatty acid degradation was decreased in an MgCl2 dose-dependent manner (Fig 4G), while the expression of genes for exogenous fatty acid degradation was increased. Western blot data confirmed that the abundance of FadL decreased with increasing MgCl2 (Fig 4H). FadH degrades unsaturated fatty acids, and influences the ratio of unsaturated to saturated fatty acids. In general, these results suggest that MgCl2 inhibits degradation of fatty acids.

FabR, FadR, ArcA, and cAMP/CRP are transcriptional factors in regulating fatty acid biosynthesis. FabR inhibits biosynthesis of unsaturated fatty acids; FadR promotes biosynthesis and inhibits fatty acids degradation; ArcA and cAMP/CRP inhibit and promote fatty acid degradation, respectively (Feng & Cronan, 2010; Fujita et al, 2007). The expression of fabR and fadR was increased, arcA was decreased. The expression of N646_1004 did not change with increasing MgCl2, while expression of N646_1885 increased only in the presence of 200 mM MgCl2 (Fig 4I). Western blot data confirmed the decreased expression of FadR with increasing MgCl2 concentration (Fig 4J). Viability of ΔfadR and ΔarcA was decreased in the presence of MgCl2, balofloxacin or both in a dose-dependent manner (Fig 4K). Thus, FadR and ArcA could play roles in bacterial resistance to balofloxacin, presumably by a mechanism that is linked to the abundance of saturated/unsaturated fatty acids.

Furthermore, to investigate whether Mg2+ functions through FadR and ArcA, expression of 8 genes of fatty acid biosynthetic genes (accA, accC, fabD, fabH, fabR, fabG, fabV, and yciA) and 5 genes of fatty acid degradation (fadL, fadD, fadB, and fadA) were quantified in wildtype, ΔfadR and ΔarcA bacterial strains. As expected, the expression of fatty acid biosynthetic genes was decreased in ΔfadR cells (Fig 4L). However, higher expression of fadD, fadB, and fadA (elevated fadA only for ΔfadR) was detected in ΔfadR or ΔacrA strains, but the expression of these genes was negatively regulated by 200 mM MgCl2 (Fig 4M-O). These results indicate that Mg2+ promotes biosynthesis of fatty acids by positively regulating fadR, and inhibits degradation of fatty acids through fadR, arcA and their downstream targets.

MgCl2 affects 16-carbon and 18-carbon fatty acid metabolism

To investigate the effect of Mg2+ on biosynthesis of saturated and unsaturated fatty acids, LC-MS was used to quantify 16-carbon and 18-carbon fatty acids, the main precursors for lipid biosynthesis (Zhang & Rock, 2008b). The abundance of saturated fatty acids, palmitic acid (C16:0), was increased, while the abundance of five unsaturated fatty acids, palmitoleic acid (C16:1), linoelaidic aid (C18:2), linoleic acid (C18:2), alpha-linoleic acid (C18:3), stearidonic acid (C18:4) was decreased with increasing MgCl2 concentration (Fig 5A). The abundance of stearic acid (C16:0) and vaccenic acid (C18:1) was unaffected. Total saturated fatty acids increased and totalunsaturated fatty acids decreased with increasing Mg2+ (Fig 5B). These results support the conclusion that Mg2+ upregulates biosynthesis of saturated fatty acids but downregulates biosynthesis of unsaturated fatty acids.

LC-MS targeted 16-carbon and 18-carbon fatty acids and role of palmitic acids and linolenic acids in BLFX resistance.

A. Scatter plots of 16-carbon and 18-carbon fatty acids, detected by LC-MS. B. Scatter plots for total saturated fatty acid and unsaturated fatty acid with16-carbon and 18-carbon in data (A). C. Synergy analysis for BLFX with linolenic acid for ATCC 33787. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (while) and non-synergistic (i.e., additive) agents (purple). D. LC- MS for abundance of intracellular linolenic acid and palmitic acid of ATCC 33787 in synergy of the indicated exogenous linolenic acid and palmitic acid. E. Synergy analysis for linolenic acid with palmitic acid for BLFX-mediated killing to ATCC 33787. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (blue) and non-synergistic (i.e., additive) agents (red). F. Bliss analysis in data (E).

Furthermore, direct exposure to palmitic acid inhibited balofloxacin-mediated killing (Fig 3F) but linolenic acid promoted balofloxacin-mediated killing in a dose-dependent manner (Fig 5C). Increasing exogenous palmitic acid and linolenic acid increased intracellular palmitic acid and linolenic acid, respectively (Fig 5D). When cells were co-treated with palmitic acid, linolenic acid, and balofloxacin, the effects of the two fatty acids appeared to be antagonistic as demonstrated by Bliss model (Fig 5E & 5F). These results indicate that saturated and unsaturated fatty acids have antagonistic effects on balofloxacin resistance.

Mg2+ promotes phospholipid biosynthesis

Fatty acids are biosynthetic precursors of lysophosphatidic acid (LPA) and phosphatidic acid (PA), two key components of the cell membrane (Zhang & Rock, 2008a). Thus, LC-MS was used to profile the effect of Mg2+ on membrane lipid composition (Fig EV6A). Increasing concentration of Mg2+ increased the percentage of lipids increased from 61% to 67%, and saturated fatty acids from 24% to 26%, but decreased the percentage of unsaturated fatty acids from 15% to 8% (Fig EV6B). The abundance of 11 lipids, 32 lipids, and 53 were increased and of 26 lipids, 52 lipids, and 107 lipids were decreased in 3.125 mM, 50 mM, and 200 mM MgCl2, respectively (Fig 6C). Saturated fatty acid and lipid increased and unsaturated fatty acid decreased in an Mg2+ dose-dependent manner (Fig EV6D). A total of 52 lipids were quantified, including 17 high abundance lipids (Fig EV6E). Phosphatidylethanolamine (PE) and phosphatidylglycerol (PG) were the first and second highest lipids in abundance (Fig EV6E) and PE decreased while PG increased with increasing Mg2+ (Figs 6A, 6B). Similar effect was observed for unsaturated derivatives of PE and PG (Fig 6C). Principal component analysis (PCA) showed that component t [1] differentiates 0 and 3.125 mM Mg2+ from 50 and 200 mM Mg2+, while component t [2] separates 50 mM Mg2+ from 0 and 200 mM Mg2+ and variants in 3.125 mM of Mg2+ treatment (Fig 6D).

Effect of Mg2+ on phospholipid biosynthesis.

A. Heatmap showing differential abundance of lipid. B. Abundance of ATCC33787 phosphatidylglycerol (PG) and phosphatidylethanolamine (PE) in the indicated concentrations of MgCl2. C. Scatter plots of PG and PE with different saturation in the presence of the indicated MgCl2. D. PCA analysis of different concentrations of MgCl2-induced phosphoglyceride metabolomes. Each dot represents the technical replicate of samples in the plot. E. S-plot generated from OPLS-DA. Predictive component p [1] and correlation p (corr) [1] differentiate 0 and 3.125 mM MgCl2 from 50 and 200 mM MgCl2. F. Scatter plot of biomarkers in data (E). G. Diagram showing phosphoglyceride metabolism. H. qRT-PCR for expression of genes encoding phosphoglyceride metabolism in the absence or the indicated concentrations of MgCl2. I. Activity of PGS and PSS in the absence or presence of the indicated concentrations of MgCl2. J. Activity of recombinant PSS in the absence or presence of the indicated concentrations of MgCl2. K and L Synergy analysis for MgCl2 with BLFX for ΔplsB (K) and ΔpgpA (L). Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (blue) and non-synergistic (i.e., additive) agents (red).

S-plot analysis showed that upregulated lipids include behenic acid, PG[16:1(9Z)/16:1(9Z)], PG[18:2 (9Z, 12Z)/16:0], PA [14:0/22:6 (4Z, 7Z, 10Z, 13Z, 16Z, 19Z)], lysoPE(16:0/0:0) and downregulated lipids include linoelaidic acid, palmitoleic acid, PE-NMe(24:0/24:0), PE[16:0/14:1(9Z)], lysoPA(i-12:0/0:0) (Fig 6E). Among them, the abundance of PG[16:1(9Z)/16:1(9Z)], PG[18:2(9Z,12Z)/16:0], and behenic acid was increased, while that of linoelaidic acid, palmitoleic acid and PE[16:0/14:1(9Z)] were decreased with increasing Mg2+ (Fig 6F). Therefore, altered phospholipid abundance may play a role in the effect of Mg2+ on resistance to balofloxacin.

PE and PG are the two end products of phospholipid metabolism (Fig 6G). Expression of most genes in the phospholipid biosynthetic pathway, including psd, pldB, and glpQ, was increased in the presence of 200 mM MgCl2, but not pssA and etuB, etuC, encoding the first enzyme and the last enzyme in the pathway, respectively. In addition, the expression of gpsA, encoding GpsA that transforms sn-glycerol-3P to glycerone-P, was increased in an independent of Mg2+ dose (Fig 6H). Phosphatidylglycerol phosphate synthase (PGS) encoded by pgsA and phosphatidylserine synthase (PSS) encoded by pssA are crucial enzymes for biosynthesis of PG and PE, respectively. Activities of PGS and PSS were elevated and reduced, respectively, by the increasing concentration of Mg2+ (Fig 6I). The effect of Mg2+ on the activity of PSS was confirmed with recombinant PSS (Fig 6J). The deletion of plsB and pgpA, the first and last genes for PG biosynthesis, respectively, lowered cell viability in the presence of balofloxacin (we failed to obtain Δpsd) (Figs 6K-I). These results indicate that phospholipids may play a role in the effect of Mg2+ on resistance to balofloxacin; more specifically, PE may inhibit and PG may promote resistance to balofloxacin.

Mg2+ regulates membrane polarization, permeability and fluidity to confer balofloxacin resistance

Membrane potential and membrane polarization influence membrane permeability and uptake of antibiotics (Lee et al., 2019; Peng et al., 2015). Thus, we speculated that exogenous Mg2+ reduces membrane polarization. Here, we show that 12.5-200 mM MgCl2 promotes membrane depolarization in a dose-dependent manner (Fig 7A). The effect of Mg2+ on proton motive force (PMF) displayed MgCl2 dose- and time- dependent manner (Fig 7B). Specifically, membrane depolarization decreased over time without MgCl2. However, depolarization gradually increased for 2 h, reached a plateau for 3 - 5 h and decreased from 5 to 10 h when supplemented with 0.78 or 3.125 mM MgCl2. The extent of maximum depolarization and the length of time at higher level of depolarization also increased by MgCl2 (Figs 7C and 7D). These results suggest that Mg2+ -dependent changes in membrane depolarization may influence bacterial resistance to antibiotics.

Mg2+ regulates membrane polarization, permeability and fluidity to confer balofloxacin resistance.

A. and B. Depolarization (A) and PMF (B) of ATCC33787 in the absence or indicated concentrations of MgCl2. C and D. Dynamic depolarization (D) Membrane fluidity of ATCC33787 in the absence or presence of the indicated concentrations of MgCl2, detected in fluorescence microscopy. E. Membrane permeability of ATCC33787 in the absence or presence of the indicated concentrations of MgCl2. F and G. Membrane permeability of ATCC33787 cultured in palmitic acid (F) or linolenic acid (G) in the indicated concentrations of MgCl2. H and I. Membrane permeability of ΔfadR (H) and ΔarcA (I) in the absence or presence of the indicated concentrations of MgCl2. J. Membrane permeability of ΔpgpA in the absence or presence of the indicated concentrations of MgCl2. K. Intracellular BLFX of ATCC33787 in the presence of palmitic acid, linolenic acid, PG, and PE (lefe panel) or in ΔfadR and ΔpgpA mutants (right panel). L. Diagram for mechanisms by which Mg2+- mediated resistance to BLFX.

Moreover, efficacy of antibiotics is strongly influenced by bacterial membrane permeability and fluidity (Saeloh et al, 2018; Zhao et al., 2021). Hence, membrane permeability and fluidity were investigated with fluorescence microscopy to visualize V. alginolyticus cells in the presence of FM5-95 (red dye) and various concentrations of MgCl2. Mg2+ decreased red staining, and no staining was observed in the presence of 200 mM Mg2+ (Fig 7D). Consistently, the uptake of SYTO9 dye decreased with increasing MgCl2, indicating bacterial membrane permeability decreased in an Mg2+dose-dependent manner (Fig 7E). These results indicate that exogenous Mg2+ decreased bacterial membrane permeability.

Furthermore, exogenous palmitic acid shifted the fluorescence signal peaks to the left in an MgCl2- and palmitic acid concentration-dependent manner (Fig 7F). In contrast, exogenous linolenic acid shifted the peak to the right in a dose-dependent manner (Fig 7G). These results support the conclusion that exogenous palmitic acid and linolenic acid decrease or increase membrane permeability, respectively. In ΔfadR cells, membrane permeability increased and then slowly decreased with increasing Mg2+, remaining higher than the control (Fig 7H). However, ΔarcA cells displayed reduced membrane permeability in the presence of low but not high MgCl2 (Fig 7I). ΔpgpA cells exhibited higher membrane permeability than control cells at all MgCl2 concentrations, but the effect was smaller at higher MgCl2 (Fig 7J). These data are consistent with the viability of these mutants as described above (Figs 4L and 6L).

Relative membrane permeability appeared to correlate with relative intracellular balofloxacin and antibiotic efficacy. Exogenous palmitic acid and linolenic acid reduced and promoted uptake of balofloxacin 159- and 18-fold, respectively (Fig 7K). Loss of fadR and pgpA increased intracellular balofloxacin (Fig 7K). These data suggest that bacterial membrane permeability is a critical factor in the efficacy and uptake of balofloxacin, in the presence or absence of exogenous MgCl2.

Discussion

The present study explores the effect of Mg2+ on phenotypic in V. alginolyticus. The results suggest that exogenous Mg2+ promotes biosynthesis of saturated fatty acids and inhibits biosynthesis of unsaturated fatty acids, while it also upregulates the activity of PGS and the abundance of PG, and downregulates the activity of PSS and abundance of PE. As a result, membrane permeability and antibiotic uptake decrease (Fig 7L), enabling phenotypic resistance. The results are consistent with the observed impact of exogenous Mg2+ on bacterial survival in the presence of balofloxacin and other antibiotics, and it is well recognized that membrane permeability critically impacts drug efficacy. Another pathway that influences drug uptake involves glutamine-inosine transport mediated by CpxA/R-OmpF (Zhao et al., 2021). The results also show that exogenous fatty acids influence membrane permeability and bacterial survival in the presence of balofloxacin.

Mg2+ is the most abundant divalent cation in the cell (Pohland & Schneider, 2019; Pontes et al, 2015) and it plays many essential roles. For example, Mg2+ stabilizes macromolecular complexes and membranes, binds cytoplasmic nucleic acids and nucleotides, interacts with phospholipid head groups and cell surface molecules, and is an essential cofactor in many enzymatic reactions (Groisman et al, 2013). A novel observation of the present study is that exogenous Mg2+ influences the abundance of palmitic acid and linolenic acid and that it upregulates PGS and downregulates PSS. These findings extend our understanding of the complexity of the biological functions of the divalent magnesium cation.

Recent studies suggest a link between fatty acid biosynthesis and antibiotic resistance in E. tarda (Su et al, 2021). The present study shows that Mg2+ has opposite effects on the abundance of saturated and unsaturated fatty acids, stimulating biosynthesis of saturated fatty acids at moderately high levels, while inhibiting biosynthesis of unsaturated fatty acids at ≥ 3 mM Mg2+. Therefore, ratio of saturated and unsaturated fatty acids should be a key clue to understand antibiotic resistance.

In previous studies of gram-positive and gran-negative bacteria (Kumariya et al, 2015; Said et al, 1987), it was shown that 10 - 20 mM Mg2+ disrupts Staphylococcus aureus membranes and kills stationary-phase S. aureus cells, but it does influence survival of Escherichia coli and Bacillus subtilis (Xie & Yang, 2016). Low concentrations of Mg2+ (≤ 10 mM) induce PmrAB-dependent modification of lipid A in wild-type E. coli (Herrera et al, 2010). A putative zwitterionic amino-containing phospholipid increased significantly, whereas amounts of phosphatidylglycerol and cardiolipin decreased in two mutants of Enterococcus faecium resistant to mundticin KS (Sakayori et al, 2003). However, Mg2+ induces antibiotic resistance through regulating the activity of PGS and PSS is not reported. More importantly, the regulation targets the two enzymes in a reverse manner to enlarge the change in the ratio between PE and PG for recognition of antibiotic resistance, which is a previously unknown mechanism mediated by Mg2+. Previous studies on membrane permeability also show that PhoP/Q is activated by low magnesium, and that inhibition or inactivation of PhoP/Q stimulates uptake of β-lactam antibiotics in Stenotrophomonas maltophilia (Huang et al, 2021). The effects of the outer membrane permeabilizers, polymyxin B nonapeptide and EDTA, are completely abolished by 3 mM Mg2+ (Kwon & Lu, 2006). In response to Mg2+ limited growth, enteric Gram-negative bacteria show higher acylation of lipid A, which alters membrane permeability and reduces uptake of cationic antimicrobial peptides (Guo et al, 1998). Mg2+ reverts the high sensitivity of rough mutants of Salmonella typhimurium, S. minnesota, and Escherichia coli 08 (i.e. with defects in the carbohydrate core of the lipopolysaccharide) to several antibiotics (Stan-Lotter et al, 1979). Mg2+ (1 mM) totally inhibited aminoglycoside-mediated outer membrane permeabilization in Pseudomonas aeruginosa (Hancock et al, 1981). However, mechanisms underlying Mg2+ regulation to antibiotic resistance are largely unknown. The present study reveals that Mg2+ plays the crucial role in the membrane permeability responsible for antibiotic resistance via modulating PE and PG in phospholipid metabolism.

In summary, the present study shows that Mg2+ modulates the phenotypic resistance of V. alginolyticus to balofloxacin and potentially other quinolones and other classes of antibiotics by a mechanism that involves altered membrane permeability and reduced antibiotic uptake. These findings inform the ongoing and future search for ways to improve the therapeutic effects of antibiotics to magnetism-induced phenotypic resistance.

Materials and Methods

Bacterial strains and culture

V. alginolyticus ATCC33787 and V. parahaemolyticus 01 are from our laboratory collection (Jiang et al, 2022; Kou et al, 2022). They were cultured in 0.5% yeast broth (HuanKai Microbial, Guangdong, China) (pH 7.0) with 3% NaCl at 30 °C overnight. The cultures were diluted 1:50 (v/v) and grown in fresh 0.5% yeast broth supplemented with desired concentrations of MgCl2 at 30 °C until OD600 = 0.6. Bacteria were harvested by centrifugation at 8,000 rpm for 3 min and resuspended in corresponding concentrations of MgCl2 to 0.6 of OD600.

Measurement of the minimum inhibitory concentration (MIC) by microtitre- dilution-method

MIC using microtitre-dilution-method was performed as previously described (Zhang et al, 2019). In brief, the overnight bacterial cultures in 3% NaCl were diluted at 1:100 (v/v) in fresh 0.5% yeast broth supplemented with desired concentrations of MgCl2, cultured at 30 °C and collected when the cells arrived at an OD600 of 0.6 in medium without MgCl2. Then, 1 X 105 CFU cells were dispensed into each well of a 96-well microtiter polystyrene tray after which a series of 2-fold dilutions of antibiotic was added. Following incubation at 30 °C for 16 h, the MIC was defined as the lowest antibiotic concentration that inhibited visible growth. Three biological repeats were carried out.

Measurement of MIC by Oxford method

Determination of MIC by Oxford cup was performed as previously described (Liu et al, 2015). Bacterial cells cultured overnight in yeast medium were diluted at 1:100, shaken at 30 0C and 200 rpm until the OD600 nm reached at 0.6. Aliquot of 100 μL cells were spread on 0.5% yeast solid medium containing 3% NaCl and 0, 0.625, 1.25, 2.5, 5, 10, 20, or 40 mM of MgCl2. Oxford cups were placed on the solid medium, and various amounts of balofloxacin (0, 0.39, 0.78, 1.56, 3.125, 6.25, and 12.5 μg) were added. After culturing at 30 °C for 12 h, diameter of the inhibition zone was measured and the inhibition area was calculated.

Measurement of intracellular balofloxacin

Measurement of intracellular balofloxacin was performed by LC-MS analysis and plate-counting assay. For LC-MS analysis, ATCC33787 cultured in the desired concentrations of MgCl2 were collected and adjusted to OD 0.6. Aliquot of 50 mL bacterial cells were added into a 250 mL Erlenmeyer flask and then 300 μL of 10 mg/mL balofloxacin (final concentration is 60 μg/mL). After being cultured for 6 h at 30 °C, these bacterial cells were collected and adjusted to OD 1.0 with 3% NaCl containing corresponding concentrations of MgCl2. Aliquot of 30 mL bacteria were collected and washed three times using mobile phase (acetonitrile: double distilled water containing 0.1mol/l formic acid = 35:65) for LC-MS detection. The bacterial cells were added 1 mL mobile phase, crushed in ice water bath for 10 min (crush for 2S, pause for 3S at 35% power). Following by centrifugation, supernatants were collected, filtered, and then measured by using liquid chromatography for balofloxacin. Different concentrations of balofloxacin were used for a standard curve. For plate-counting assay, ATCC33787 cells cultured overnight were diluted at 1:100, shaken at 30 0C and 200 rpm until the OD600 nm reached at 0.6. The cells were diluted at 1:100 and spread on LB agar to preparing ATCC33787 plates. Meanwhile, ATCC33787 cells were cultured in medium with desired MgCl2 and at 30 °C and 200 rpm for 6 h. These cells were collected, washed and sonicated. The sonicated solution was added to Oxford cups with the ATCC33787 plates. After culturing at 30 °C for 12 h, diameter of the inhibition zone was measured. A gradient of balofloxacin was used as a standard curve for drug quantification.

Measurement of intracellular magnesium

Measurements of intracellular Mg2+ concentration were carried out as previously described with a modification (Yang et al., 2018). In brief, V. alginolyticus cultured in 0%, 0.78%, 3.125%, 12.5%, 50%, and 200% MgCl2 were collected, washed, and resuspended in the same concentrations of MgCl2 to an OD600 of 1.0. Aliquots of 20 mL of bacterial suspensions were centrifuged and washed by ddH2O once. The resulting cells were weighted, which was designated wet weight. These samples were freeze-dried overnight, which was designated dry weight. Intracellular water volume was calculated using the following formula: W × (1-W/D-0.23) as described previously (Unemoto et al, 1973). Two hundred microliters of concentrated nitric acid was added to the dried cells and then heated in 75 °C for 20 min. After diluting the samples 20- fold, they were analyzed for Mg2+ concentration by inductively coupled plasma mass spectrometry (ICP-MS) (iCAP 6500, Thermo Fisher). The intracellular Mg2+ concentration was calculated according to the volume of intracellular water.

Western blotting

Western blotting was carried out as described previously (Yao et al, 2019). Bacterial protein samples were prepared by ultrasound treatment, resolved on a 12% SDS-PAGE and transferred to nitrocellulose membranes (GE Healthcare Life Sciences). The membranes were incubated with 1:100 of the primary mouse antibodies, followed by goat anti-mouse secondary antibodies conjugated with horseradish peroxidase. Band intensities were detected by using a chemiluminescence imaging analysis system, Tanon-5200.

Metabolomics analysis

Metabolomics analysis was performed by GC-MS as described previously (Yang et al., 2018). Briefly, ATCC33787 were cultured in 0.5% yeast broth with desired MgCl2. Equivalent numbers of cells were quenched with 60% (v/v) cold methanol (Sigma) and then centrifuged at 8,000 rpm at 4 °C for 5 min. One milliliter of cold methanol was used to extract metabolites. To do this, the samples were sonicated for 5 min at a 10- Wpower setting using the Ultrasonic Processor (JY92-IIDN, Scientz, China), followed by centrifugation at 12,000 rpm in 4 °C for 10 min. Supernatants were collected and 10 µL ribitol (0.1 mg per mL, Sigma-Aldrich, USA) was added into each sample as an internal quantitative standard. The supernatants were concentrated for metabolite derivatization and then used for GC-MS analysis. Every experiment was repeated by five biological replicates. GC-MS detection and spectral processing for GC-MS were carried out using the Agilent 7890A GC equipped with an Agilent 5975C VL MSD detector (Agilent Technologies, USA) as described previously (Unemoto et al., 1973; Yang et al., 2018). Statistical difference was obtained by Kruskal Wallis test and Manne Whitney test using SPSS 13.0 and a p value < 0.01 was considered significant.

Hierarchical clustering was completed in the R platform (https://cran.r-project.org/) with the function “heatmap. 2” of “gplots library”. Z score analysis was used to scale each metabolite. Multivariate statistical analysis included principal component analysis (PCA) and orthogonal partial least square discriminant analysis (OPLS-DA) implemented with SIMCA 12.0 (Umetrics, Umeå, Sweden). Control scaling was selected prior to fitting. All variables were mean centered and scaled to pareto variance of each variable. PCA was used to reduce the high dimension of the data set. Differential metabolites to their respective biochemical pathways were outlined in the MetaboAnalyst 3.0 (http://www.metaboanalyst.ca/). Pathways were enriched by raw p value < 0.05.

qRT-PCR

Quantitative real time polymerase chain reaction (qRT-PCR) was carried out as described previously (Yang et al, 2020). Total RNA was extracted from V. alginolyticus using TRIZOL regent (Invitrogen Life Technologies) according to the protocol. RNA Electrophoresis was carried out in 1% (w/v) agarose gels to identify quality of the extracted RNA. By using a PrimeScriptTM RT reagent Kit with gDNA eraser (Takara, Japan), reverse transcription-PCR was carried out on 1 μg of total RNA and primers are listed in Table EV2. qRT-PCR was performed in 384-well plates with a total volume of 10 μL and the reaction mixtures were run on a LightCycler 480 system (Roche, Germany). Data are shown as the relative mRNA expression compared to 0% MgCl2 test with the endogenous reference 16S rRNA gene.

Antibiotic bactericidal assay

Antibiotic bactericidal assay was performed as described previously with a modification (Li et al, 2016; Zhao et al., 2021). The cultured bacteria of ATCC33787 and its mutant strains were transferred to fresh yeast medium at a dilution of 1:1000 and dispensed into test tubes and then the indicated concentrations of MgCl2 and balofloxacin were added. If desired, 2 mM 2-aminooxazole and 1 μg/mL triclosan were or a metabolite complemented. These mixtures were incubated at 30 °C and 200 rpm for 6 h. Cells were collected. To determine CFU per mL, 100-μL samples were 10-fold serially diluted with 900 μL M9 buffer and an aliquot of 5 μL of each dilution was spotted onto the LB agar plates and cultured at 30 °C for 8 h.

Construction of gene-deleted mutants

Construction of gene-deleted mutants was carried out as described previously (Kuang et al, 2021). Primers were designed as shown in Table EV3 using CE Design V1.03 software. To construct gene-deleted mutants, upstream and downstream 500-bp fragments were amplified from the genome using two pairs of primers (primers P1 and P2, primers P3 and P4), and then merged into a 1,000-bp fragment by overlap PCR using a pair of primers (primers P1 and P4). After the fragments were digested by XbaI, they were ligated into the pDS132 vector digested by the same enzymes, transformed into MC1061 competent cells. Conversion products were coated on LB plate containing 25 μg/mL chloramphenicol and cultured at 37 ℃ overnight. The plasmids from colony growing on the plate were identified by PCR using a pair of primers (primers P1 and P4) and sequenced. The sequenced plasmids were transformed into MFD λ pir competent cells as donor. MFD λpir and recipient bacterium ATCC 33787 were cultured to an optical density (OD) of 1.0 and then mixed at a ratio of 4:1. After centrifugation, the pellets were resuspended with 50 μL LB medium including DAP (100 μg/mL), dropped onto sterilized filter paper on LB medium including DAP (100 μg/mL), and cultured for 16-18 h at 37 °C.

All bacteria rinsed from the filter paper with LB medium were smeared onto the LB plate with chloramphenicol (25 μg/mL) and Ampicillin (100 μg/mL). After cultured for 16-18 h at 37 °C, bacteria were screened by LB plate with above two antibiotics. The bacteria were identified by plasmid PCR using a pair of primers P1 and P4 and sequencing and then were cultured and smeared onto the LB plates with 20% sucrose. The clones were cultured and smeared onto the LB plates with 20% sucrose again. The clones, which did not grow on the LB plates with chloramphenicol but grew on the LB plates with 20% sucrose, were identified by PCR using primer P7P8 (primer P7 is set at about 250 bp upstream of P1, and primer P8 is set at about 250 bp downstream of P4), P4P7, and P5P6 (for amplification of the target gene).

LC-MS analysis for lipidomics

ATCC33787 were cultured in medium with desired concentrations of MgCl2, washed by 3% NaCl with the desired concentrations of MgCl2, and adjusted to OD 1.0. Aliquot of 30 mL the cultures was centrifuged each sample and bacterial cells were collected. Aliquot of 800 μL distilled water was added and treated for 5 min at 100 °C to inactivate phospholipase C. Protein concentration was determined with BCA kit. Sample solution with 3 mg protein was transferred into a 10 mL centrifuge tube, and supplemented with distilled water to 800 μL. Then all samples were processed as follows: 1 mL chloroform and 2 mL methanol (to make the volume ratio of chloroform: methanol: water = 1:2 : 0.8) were added and cholic acid was as an internal standard; These mixtures were vortexed for 2 min and then 1 mL chloroform was add and vortex for 30s; 1 mL 10% NaCl solution was add, vortexed for 30s, and placed at room temperature overnight. After the solution was layered, a 2.5 mL syringe was used to suck the lowest layer (chloroform layer) to a new EP tube. Solution of the chloroform layer was evaporated in the rotary evaporator for 2 h and then 1 mL mobile phase (50% and 50%B. A, 25 mM ammonium acetate / methanol (30 : 70); B, methanol) was added for analysis of lipids by LC-MS. Database (https://hmdb.ca) was used for lipid identification.

Measurement of enzyme activity

Enzyme activity was carried out as previously described (Jiang et al, 2020). Cells cultured in medium were collected and washed three times with saline. The bacterial cells were suspended in Tris–HCl (pH 7.4) and disrupted by sonic oscillation for 6 min (200 W total power with 35% output, 2 s pulse, 3 s pause over ice). After centrifugation, supernatants were collected. The protein concentration of the supernatant was determined using Bradford assay (Beyotime, P0009). Then, 200 μg proteins were used for determination of pyruvate dehydrogenase (PDH), α-ketoglutarate dehydrogenase (KGDH) and succinate dehydrogenase (SDH) activity.

Measurement of proton motive force

Measurement of the transmembrane voltage delta PSI (δPSI), which is the electrical component of the PMF, was performed as previously described (Cheng et al, 2018). Bacteria cultured in medium with desired concentrations of MgCl2 were collected at centrifugation and labeled by DiO2(3). Approximately 1 × 107 CFU were added to a flow cytometry tube containing 1 mL buffer with 10 μM DiOC2(3) (Sigma) and incubated in the dark for 30 min at 30 0C. Samples were assayed with BD FACSCalibur flow cytometer with a 488 nm excitation wavelength. Gates for bacterial populations were based on the control population by using forward versus side scatter and red versus green emission. Size and membrane potential determined the intensity of Red (488 nm excitation, 610 nm emission) fluorescence. The diverse ratios of red and green indicated fluorescence intensity values of the gated populations. Computational formula of membrane potential:

Measurement of depolarization

The membrane potential was estimated by measuring the fluorescence of the potential- dependent probe DiBAC4(3) (Saint-Ruf et al, 2016). ATCC33787 were cultured in medium with desired concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacterial cells were diluted to 1 mL, and 2 μL of 5 mM DiBAC4(3) was added. After incubated for 15 min at 30 ℃ without light and vibration, these samples were filtered and detected by BD FACSCalibur.

Measurement of fluidity by fluorescence microscopy

ATCC33787 were cultured in medium with desired concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacteria cells of each sample were diluted to 1 mL and 10 μL (10 mg/mL) FM5-95 was added. After incubated for 20 min at 30 ℃ at vibration without light, the sample was centrifuged for 10 min at 12,000 rpm. The pellets were resuspended with 20 μL of 3% NaCI. Aliquot of 2 μL sample was dropped on the agarose slide, and take photos under the inverted fluorescence microscope.

Measurement of membrane permeability

Measurement of membrane permeability was carried out as described previously (Su et al., 2021). ATCC33787 were cultured in medium with desired concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacteria cells of each sample were diluted to 1 mL and 2 μL 10 mg/mL SYT09 was added. After incubated for 15 min at 30 ℃ at vibration without light, the mixtures were filtered and measured by BD FACSCalibur.

eLife Assessment

The study explored the influence of magnesium on phenotypic antibiotic resistance in two Vibrio model bacteria. This research is fundamental for revealing the phenotypic antibiotic resistance mechanism utilized by the specified model bacteria in elevated levels of magnesium. The study produced solid evidence indicating that in high concentrations of magnesium, the efficacy of selected antibiotics was diminished due to decreased biosynthesis of unsaturated fatty acids and PE, along with an increase in the biosynthesis of PG.

Reviewer #1 (Public Review):

Summary:

In the manuscript entitled "Magnesium modulates phospholipid metabolism to promote bacterial phenotypic resistance to antibiotics", Li et al demonstrated the role of magnesium in promoting phenotypic resistance in V. alginolyticus. Using standard microbiological and metabolomic techniques, the authors have shown the significance of fatty acid biosynthesis pathway behind the resistance mechanism. This study is significant as it sheds light on the role of an exogenous factor in altering membrane composition, polarization, and fluidity which ultimately leads to antimicrobial resistance.

Strengths:

(1) The experiments were carried out methodically and logically.

(2) An adequate number of replicates were used for the experiments.

Weaknesses:

(1) The introduction section needs to be more informative and to the point.

(2) The weakest point of this paper is in the logistics through the results section. The way authors represented the figures and interpreted them in the results section (or the figure legends) does not match. The figures are difficult to interpret and are not at all self-explanatory.

(3) There are too many mislabeling of the figure panels in the main text which makes it difficult to find out which figures the authors are explaining. There should be more explanation on why and how they did the experiments and how the results were interpreted.

Reviewer #2 (Public Review):

Summary:

In this study, the authors aimed to identify if and how magnesium affects the ability of two particular bacteria species to resist the action of antibiotics. In my view, the authors succeeded in their goals and presented a compelling study that will have important implications for the antibiotic resistance research community. Since metals like magnesium are present in all lab media compositions and are present in the host, the data presented in this study certainly will inspire additional research by the community. These could include research into whether other types of metals also induce multi-drug resistance, whether this phenomenon can be observed in other bacterial species, especially pathogenic species that cause clinical disease, and whether the underlying molecular determinants (i.e. enzymes) of metal-induced phenotypic resistance could be new antimicrobial drug targets themselves.

Strengths:

This study's strengths include that the authors used a variety of methodologies, all of which point to a clear effect of exogenous Mg2+ on drug resistance in the targeted species. I also commend the authors for carrying out a comprehensive study, spanning evaluation of whole cell phenotypes, metabolic pathways, genetic manipulation, to enzyme activity level evaluation. The fact that the authors uncovered a molecular mechanism underlying Mg2+-induced phenotypic resistance is particularly important as the key proteins should be studied further.

Weaknesses:

I believe there are weaknesses in the manuscript, however. The authors take for granted that the reader is familiar with all the assays utilized, and do not properly explain some experiments, and thus I highly suggest that the authors add a brief statement in each situation describing the rationale for each selected methodology (more details are in the private review to the authors). The Results section is also quite long and bogs down at times, and I suggest that the authors reduce its length by 10 to 20%. In contrast, the Introduction is sparse and lacks key aspects, for example, there should be mention of the study's main purpose and approaches, plus an introduction to the authors' choice of species and their known drug resistance properties, as well as the drug of choice (balofloxacin). Another notable weakness is that the authors evaluated Mg2+-induced phenotypic resistance only against two closely related species, and thus the generalizability of this mechanism of drug resistance is not known. The paper would be strengthened if the authors could demonstrate this type of phenotypic resistance in at least one more Gram-negative species and at least one Gram-positive species (antimicrobial susceptibility evaluations would suffice), each of which should be pathogenic to humans. Demonstrating magnesium-induced phenotypic drug resistance in the WHO Priority Bacterial Pathogens would be particularly important.

In general, the conclusions drawn by the authors are justified by the data, except for the interpretation of some experiments. Importantly, this paper has discovered new antimicrobial resistance mechanisms and has also pointed to potential new targets for antimicrobials.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public Review):

Summary:

In the manuscript entitled "Magnesium modulates phospholipid metabolism to promote bacterial phenotypic resistance to antibiotics", Li et al demonstrated the role of magnesium in promoting phenotypic resistance in V. alginolyticus. Using standard microbiological and metabolomic techniques, the authors have shown the significance of fatty acid biosynthesis pathway behind the resistance mechanism. This study is significant as it sheds light on the role of an exogenous factor in altering membrane composition, polarization, and fluidity which ultimately leads to antimicrobial resistance.

Strengths:

(1) The experiments were carried out methodically and logically.

(2) An adequate number of replicates were used for the experiments.

Weaknesses:

(1) The introduction section needs to be more informative and to the point.

Thank you so much for your suggestion. We have revised the introduction to make it more informative and to the point as following:

“Non-inheritable antibiotic or phenotypic resistance represents a serious challenge for treating bacterial infections. Phenotypic resistance does not involve genetic mutations Phenotypic resistance does not involve genetic mutations and is transient, allowing bacteria to resume normal growth. Biofilm and bacterial persisters are two phenotypic resistance types that have been extensively studied (Brandis et al., 2023; Corona & Martinez, 2013). Biofilms have complex structures, containing elements that impede antibiotic diffusion, sequestering and inhibiting their activity (Ciofu et al., 2022). Biofilm-forming bacteria and persisters also have distinct metabolic states that significantly reduce their antibiotic susceptibility (Yan & Bassler, 2019). These two types of phenotypic resistance share the common feature in their retarded or even cease of growth in the presence of antibiotics (Corona & Martinez, 2013). However, specific factors that promote phenotypic resistance and allow bacteria to proliferate in the presence of antibiotics remain poorly defined.

Metal ions have a diverse impact on the chemical, physical, and physiological processes of antibiotic resistance (Booth et al, 2011; Lu et al, 2020; Poole, 2017). This includes genetic elements that confer resistance to metals and antibiotics (Poole, 2017) and metal cations that directly hinder (or enhance) the activity of specific antibiotic drugs (Zhang et al., 2014). The metabolic environment can also impact the sensitivity of bacteria to antibiotics (Jiang et al., 2023; Lee & Collins, 2012; Peng et al., 2015; Zhang et al., 2020; Zhao et al., 2021). Light metal ions, such as magnesium, sodium, and potassium, can behave as cofactors for different enzymes (Du et al., 2016) and influence drug efficacy. Heavy metal ions, including Cu2+ and Zn2+, confer resistance to antibiotics (Yazdankhah et al., 2014; Zhang et al., 2018). Recent reports suggest that sodium negatively regulates redox states to promote the antibiotic resistance of Vibrio alginolyticus (Yang et al., 2018), while actively growing Bacillus subtilis cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al, 2019). In Gram-negative bacteria, by contrast, zinc enhances antibiotic efficacy by potentiating carbapenem, fluoroquinolone, and β-lactam-mediated killing (Isaei et al., 2016; Zhang et al., 2014). Magnesium influences bacterial structure, cell motility, enzyme function, cell signaling, and pathogenesis (Wang et al., 2019). This mineral also modulates microbiota to harvest energy from the diet (Garcia-Legorreta et al., 2020), allowing Bacillus subtilis to cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al., 2019). However, the role of magnesium in promoting phenotypic resistance is less well understood.

Vibrios inhabit seawater, estuaries, bays, and coastal waters, regions full of metal ions such as magnesium (Kumarage et al., 2022). Magnesium is the second most dissolved element in seawater after sodium. At a salinity of 3.5% seawater, the magnesium concentration is about 54 mM (Potis, 1968), and in deep seawater, can be as high as 2,500 mM (Wang et al., 2024). Vibrio parahaemolyticus and V. alginilyticus are two representative Vibrio pathogens that infect humans and aquatic animals, resulting in illness and economic loss, respectively (Grimes, 2020). (Fluoro)quinolones such as balofloxacin are used to treat Vibrio infection, however, resistance has emerged due to overuse (Suyamud et al., 2024). Indeed, (fluoro)quinolones are one of China's two primary residual chemicals associated with aquaculture (Liu et al., 2017). Vibrio can develop quinolone resistance through mutations in the DNA gyrase gene or through plasmid-mediated mechanisms (Dutta et al., 2021). Thus, the use of V. parahaemolyticus and V. alginilyticus as bacterial representatives, and balofloxacin as a quinolone-based antibacterial representative, can help to define novel magnesiumdependent phenotypic resistance mechanisms of pathogenic Vibrio species.

The current study evaluated whether magnesium induces phenotypic resistance in Vibrio species and defined the molecular/genetic basis for this resistance. Genetic approaches, GC-MS analysis of metabolite and membrane remodeling upon antibiotic exposure, membrane physiology, and extensive antimicrobial susceptibility testing were used for the evaluations.”

(2) The weakest point of this paper is in the logistics through the results section. The way authors represented the figures and interpreted them in the results section (or the figure legends) does not match. The figures are difficult to interpret and are not at all self-explanatory.

Thank you so much for your suggestion. We have followed your suggestion to check the match between result and figures. They are now revised.

(3) There are too many mislabeling of the figure panels in the main text which makes it difficult to find out which figures the authors are explaining. There should be more explanation on why and how they did the experiments and how the results were interpreted.

Thank you so much for your suggestion. We have checked the figures and main text to ensure that we make every figure clearly stated.

Reviewer #2 (Public Review):

Summary:

In this study, the authors aimed to identify if and how magnesium affects the ability of two particular bacteria species to resist the action of antibiotics. In my view, the authors succeeded in their goals and presented a compelling study that will have important implications for the antibiotic resistance research community. Since metals like magnesium are present in all lab media compositions and are present in the host, the data presented in this study certainly will inspire additional research by the community. These could include research into whether other types of metals also induce multi-drug resistance, whether this phenomenon can be observed in other bacterial species, especially pathogenic species that cause clinical disease, and whether the underlying molecular determinants (i.e. enzymes) of metal-induced phenotypic resistance could be new antimicrobial drug targets themselves.

Strengths:

This study's strengths include that the authors used a variety of methodologies, all of which point to a clear effect of exogenous Mg2+ on drug resistance in the targeted species. I also commend the authors for carrying out a comprehensive study, spanning evaluation of whole cell phenotypes, metabolic pathways, genetic manipulation, to enzyme activity level evaluation. The fact that the authors uncovered a molecular mechanism underlying Mg2+-induced phenotypic resistance is particularly important as the key proteins should be studied further.

Weaknesses:

I believe there are weaknesses in the manuscript, however. The authors take for granted that the reader is familiar with all the assays utilized, and do not properly explain some experiments, and thus I highly suggest that the authors add a brief statement in each situation describing the rationale for each selected methodology (more details are in the private review to the authors). The Results section is also quite long and bogs down at times, and I suggest that the authors reduce its length by 10 to 20%. In contrast, the Introduction is sparse and lacks key aspects, for example, there should be mention of the study's main purpose and approaches, plus an introduction to the authors' choice of species and their known drug resistance properties, as well as the drug of choice (balofloxacin). Another notable weakness is that the authors evaluated Mg2+-induced phenotypic resistance only against two closely related species, and thus the generalizability of this mechanism of drug resistance is not known. The paper would be strengthened if the authors could demonstrate this type of phenotypic resistance in at least one more Gram-negative species and at least one Gram-positive species (antimicrobial susceptibility evaluations would suffice), each of which should be pathogenic to humans. Demonstrating magnesium-induced phenotypic drug resistance in the WHO Priority Bacterial Pathogens would be particularly important.

In general, the conclusions drawn by the authors are justified by the data, except for the interpretation of some experiments. Importantly, this paper has discovered new antimicrobial resistance mechanisms and has also pointed to potential new targets for antimicrobials.

Thank you so much for your suggestion! We followed your idea the revise the manuscript as following:

(1) We added a brief statement in the situation to explain the result and methodology according to your suggestion in the private review.

(2) To make the streamline of the story more logic, we moved the whole second result to supplementary text and supplementary figure.

(3) We revised the introduction part by adding additional information to make it informative and to the point as following:

“Non-inheritable antibiotic or phenotypic resistance represents a serious challenge for treating bacterial infections. Phenotypic resistance does not involve genetic mutations Phenotypic resistance does not involve genetic mutations and is transient, allowing bacteria to resume normal growth. Biofilm and bacterial persisters are two phenotypic resistance types that have been extensively studied (Brandis et al., 2023; Corona & Martinez, 2013). Biofilms have complex structures, containing elements that impede antibiotic diffusion, sequestering and inhibiting their activity (Ciofu et al., 2022). Biofilm-forming bacteria and persisters also have distinct metabolic states that significantly reduce their antibiotic susceptibility (Yan & Bassler, 2019). These two types of phenotypic resistance share the common feature in their retarded or even cease of growth in the presence of antibiotics (Corona & Martinez, 2013). However, specific factors that promote phenotypic resistance and allow bacteria to proliferate in the presence of antibiotics remain poorly defined.

Metal ions have a diverse impact on the chemical, physical, and physiological processes of antibiotic resistance (Booth et al, 2011; Lu et al, 2020; Poole, 2017). This includes genetic elements that confer resistance to metals and antibiotics (Poole, 2017) and metal cations that directly hinder (or enhance) the activity of specific antibiotic drugs (Zhang et al., 2014). The metabolic environment can also impact the sensitivity of bacteria to antibiotics (Jiang et al., 2023; Lee & Collins, 2012; Peng et al., 2015; Zhang et al., 2020; Zhao et al., 2021). Light metal ions, such as magnesium, sodium, and potassium, can behave as cofactors for different enzymes (Du et al., 2016) and influence drug efficacy. Heavy metal ions, including Cu2+ and Zn2+, confer resistance to antibiotics (Yazdankhah et al., 2014; Zhang et al., 2018). Recent reports suggest that sodium negatively regulates redox states to promote the antibiotic resistance of Vibrio alginolyticus (Yang et al., 2018), while actively growing Bacillus subtilis cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al, 2019). In Gram-negative bacteria, by contrast, zinc enhances antibiotic efficacy by potentiating carbapenem, fluoroquinolone, and β-lactam-mediated killing (Isaei et al., 2016; Zhang et al., 2014). Magnesium influences bacterial structure, cell motility, enzyme function, cell signaling, and pathogenesis (Wang et al., 2019). This mineral also modulates microbiota to harvest energy from the diet (Garcia-Legorreta et al., 2020), allowing Bacillus subtilis to cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al., 2019). However, the role of magnesium in promoting phenotypic resistance is less well understood.

Vibrios inhabit seawater, estuaries, bays, and coastal waters, regions full of metal ions such as magnesium (Kumarage et al., 2022). Magnesium is the second most dissolved element in seawater after sodium. At a salinity of 3.5% seawater, the magnesium concentration is about 54 mM (Potis, 1968), and in deep seawater, can be as high as 2,500 mM (Wang et al., 2024). Vibrio parahaemolyticus and V. alginilyticus are two representative Vibrio pathogens that infect humans and aquatic animals, resulting in illness and economic loss, respectively (Grimes, 2020). (Fluoro)quinolones such as balofloxacin are used to treat Vibrio infection, however, resistance has emerged due to overuse (Suyamud et al., 2024). Indeed, (fluoro)quinolones are one of China's two primary residual chemicals associated with aquaculture (Liu et al., 2017). Vibrio can develop quinolone resistance through mutations in the DNA gyrase gene or through plasmid-mediated mechanisms (Dutta et al., 2021). Thus, the use of V. parahaemolyticus and V. alginilyticus as bacterial representatives, and balofloxacin as a quinolone-based antibacterial representative, can help to define novel magnesiumdependent phenotypic resistance mechanisms of pathogenic Vibrio species.

The current study evaluated whether magnesium induces phenotypic resistance in Vibrio species and defined the molecular/genetic basis for this resistance. Genetic approaches, GC-MS analysis of metabolite and membrane remodeling upon antibiotic exposure, membrane physiology, and extensive antimicrobial susceptibility testing were used for the evaluations.”

(4) We examined the effect of magnesium in WHO listed priority strains, which confirmed the results as following:

“Importantly, exogenous MgCl2 also increased MICs of clinic isolates, carbapenemresistant Escherichia coli, carbapenem-resistant Klebsiella pneumoniae, carbapenemresistant Pseudomonas aeruginosa and carbapenem-resistant Acinetobacter baumannii to balofloxacin (Fig 1G).”

Recommendations for the authors:

Reviewer #1 (Recommendations For The Authors):

(1) There are many grammatical mistakes to point out. The manuscript needs proofreading and editing.

We appreciate this comment! The manuscript has been revised by a native speaker.

(2) The introduction could be more informative. A little more description of magnesium - such as what it does to antibiotics and how it's known to affect the microbiome - might be helpful for the general readers. The question remains why out of all the metal ions that might affect antibiotic resistance (many of them are less explored), authors particularly decided to work on the effect of magnesium. The introduction should cover the rationale of their hypothesis. Also, the authors might want to briefly talk about the model organisms (V. algonolyticus and V. parahemolyticus) describing how threatening they are and how they are becoming resistant to antibiotics.

We appreciate this comment! We revise the introduction by providing additional information as following:

“In Gram-negative bacteria, by contrast, zinc enhances antibiotic efficacy by potentiating carbapenem, fluoroquinolone, and β-lactam-mediated killing (Isaei et al., 2016; Zhang et al., 2014). Magnesium influences bacterial structure, cell motility, enzyme function, cell signaling, and pathogenesis (Wang et al., 2019). This mineral also modulates microbiota to harvest energy from the diet (Garcia-Legorreta et al., 2020), allowing Bacillus subtilis to cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al., 2019). However, the role of magnesium in promoting phenotypic resistance is less well understood.

Vibrios inhabit seawater, estuaries, bays, and coastal waters, regions full of metal ions such as magnesium (Kumarage et al., 2022). Magnesium is the second most dissolved element in seawater after sodium. At a salinity of 3.5% seawater, the magnesium concentration is about 54 mM (Potis, 1968), and in deep seawater, can be as high as 2,500 mM (Wang et al., 2024). Vibrio parahaemolyticus and V. alginilyticus are two representative Vibrio pathogens that infect humans and aquatic animals, resulting in illness and economic loss, respectively (Grimes, 2020). (Fluoro)quinolones such as balofloxacin are used to treat Vibrio infection, however, resistance has emerged due to overuse (Suyamud et al., 2024). Indeed, (fluoro)quinolones are one of China's two primary residual chemicals associated with aquaculture (Liu et al., 2017). Vibrio can develop quinolone resistance through mutations in the DNA gyrase gene or through plasmid-mediated mechanisms (Dutta et al., 2021). Thus, the use of V. parahaemolyticus and V. alginilyticus as bacterial representatives, and balofloxacin as a quinolone-based antibacterial representative, can help to define novel magnesiumdependent phenotypic resistance mechanisms of pathogenic Vibrio species.

The current study evaluated whether magnesium induces phenotypic resistance in Vibrio species and defined the molecular/genetic basis for this resistance. Genetic approaches, GC-MS analysis of metabolite and membrane remodeling upon antibiotic exposure, membrane physiology, and extensive antimicrobial susceptibility testing were used for the evaluations. ”

(3) Figure 1C is mislabeled as 1B (line 100). Line 101: The sentence is not clear and very confusing. What is meant by 15.6mM - 62.4 mM? Are they talking about the concentration of BLFX (though in the figure the concentration was shown in µg)? Please rewrite the sentence in a simplified way. Also, the zone of inhibition was decreased with increasing MgCl2, not increased.

We appreciate this comment! These have been revised, including that Fig 1B is now corrected as Fig. 1C. Line 101, which is now Line 122. The sentence was revised as following:

“At balofloxacin doses of 1.56, 3.125, 6.25, and 12.5 µg, the zone of inhibition decreased with increasing MgCl2 (Fig 1D)”

(4) In the western blot images, it would be nice to indicate the MW of the protein bands shown. The loading control used for the experiments should be clearly mentioned in the figure legends.

We appreciate this comment! The MWs are indicated in the western-blot image throughout the manuscript.

The loading control is clearly stated in the figure legend as following:

“Whole cell lysates resolved by SDS-PAGE gel was stained with Coomassie brilliant blue as loading control.”.

(5) Figures 2 B and C: the figure legend does not explain what the authors wanted to show. It's not clear how they plotted the inhibitory curve, or the binding efficacy. These panels need an explanation of how the analysis was done.

We appreciate this comment! The figure 2 is now removed to Suppl. Fig 2, and the description of figure 2 is moved to Suppl. Text. We revise the description of the result as following, which is in Suppl. Text:

“Prior studies suggest that the chelation of antibiotics by magnesium ions inhibits antibiotic uptake (Deitchman et al., 2018; Lunestad and Goksøyr, 1990). To investigate whether magnesium binds to balofloxacin, balofloxacin was pre-incubated with magnesium, and zone of inhibition (ZOI) analysis was conducted. Six different concentrations of balofloxacin were separately incubated with six different concentrations of MgCl2, and then spotted on filter paper so that a defined amount of balofloxacin could be used for ZOI. While lower concentrations of MgCl2, (0.78, 3.125, or 12.5 mM) did not alter the ZOI, higher concentrations, including 50 and 200 mM MgCl2, decreased the ZOI (Suppl. Fig 2A), suggesting that even high doses of magnesium had only a partial effect on balofloxacin through direct binding. For example, at 200 mM MgCl2 and 5 or 10 μg/mL balofloxacin, the balofloxacin ZOI was 53.2 and 70.3% of the ZOI at 0 mM MgCl2, suggesting that 50% of the antibiotics were still functional. Intracellular BLFX also decreased with increasing MgCl2 (Suppl. Fig 2B), while exogenous Mg2+ increased intracellular Mg2+ levels in a dose-dependent manner. For example, exogenous 50 and 200 mM MgCl2 increased intracellular Mg2+ levels to 1.21 and 1.31 mM, respectively (Suppl. Fig 2C). The relationship between TolC, an efflux pump that transports quinolones from bacterial cells, and Mg2+ was also assessed (Kobylka et al., 2020; Song et al., 2020). The expression of TolC/tolC was unaffected by Mg2+ (Suppl. Fig 2D). Magnesium is critical for LPS stability. LPS levels increased at 200 mM Mg2+ (Suppl. Fig 2E), however, the loss of waaF, lpxA, and lpxC, three key genes involved in LPS biosynthesis, did not influence balofloxacin sensitivity/resistance in the presence of Mg2+ (Suppl. Fig 2F). These findings suggest that magnesium-induced LPS biosynthesis does not contribute directly to BLFX resistance and demonstrate that Mg2+ influx is involved in balofloxacin resistance.”

(6) For the metabolomics results, it will help immensely if the authors provide a volcano plot of the identified metabolites and plot the heat map according to the -log2 metabolite intensities. In Figure 3A, it's not clear what information is conveyed through Euclidean distance calculations of the heat map. In Figure 3 B, the authors mentioned that the OPLS-DA test was conducted, although the figure shows a PCA plot, so it's not clear how these two are connected. Figure 3 E: the figure legend says scattered plot, but the panel represents color-coded numerical values, not a scattered plot. Also, it's not clear how they got those values.

We appreciate this comment! We quite agree with you that if the differential metabolites could be shown as volcano plot. However, we didn’t adopt volcano plot in this study because this is a magnesium concentration-dependent metabolomes that includes 6 groups in parallel. Volcano plots may give a complex view of the comparison among different groups. We also tried to plot the heat map according to the -log2 metabolite intensities. Although this analysis cluster 200 mM and 50 mM groups better, the data of low magnesium concentrations was not consistent, which may be due to the minor metabolic change of low concentrations magnesium. Thank you for your understanding.

For Euclidean distance calculations, we explain in the figure legend as following:

“Euclidean distance calculations were used to generate a heatmap that shows clustering of the biological and technical replicates of each treatment.”

In Figure 2B, which was Figure 3B in previous version, it has been replaced with OPLS-DA analysis in the revised version.

In Figure 2E, which was Figure 3E in previous version, it is revised as following:

“E. Areas of the peaks of palmitic acid and stearic acid generated by GC-MS analysis.”

(7) In Figure 4, the figure legends (as well as the in the text) are not properly referred to. Please make sure to refer to the correct panel.

We appreciate this comment! The figure legends have been corrected to match the panel and text.

Figure 4F: how was the synergy analysis done? In the methods section, the authors described the antibiotic bactericidal assay protocol, but there was no clear indication of how they generated the isobologram.

We appreciate this comment! We provide additional information in the Figure 3F legend, which was Figure 4F in previous version, as following:

“Synergy analysis for BFLX with palmitic acid for V. alginolyticus. Synergy was performed by comparing the dose needed for 50% inhibition of the synergistic agents (white) and non-synergistic (i.e., additive) agents (purple).”

(8) Figure 5 A: the scatter plot is plotted according to the area along the Y axis: which "area" is represented here? There is absolutely no explanation, neither in the results nor in the figure legends. Using box plots might be a better option than using a scattered plot.

We appreciate this comment! “Area” has been noted in the revised manuscript as following:

“The area indicates the area of the peak of the metabolite in total ion chromatography of GC-MS.”

(9) In Figure 6 A, the heat map is plotted according to the column Z scores. What is meant by "column Z score"? The corresponding figure legend says, "heat map showing differential abundance of lipid". Z scores do not represent an abundance of a variable, so the conclusion might not be appropriate here.

We appreciate this comment! In Figure 5A, which was Figure 6A in previous version, column Z score shows the abundance of metabolites analyzed, which is automatically generated in the heat map analysis to give a sign of these metabolites tested. The legend has been revised as following:

“Heatmap showing changes in differential lipid levels at the indicated concentration of MgCl2.”

(10) Line 313-314: it should be Figure EV6C.

We appreciate this comment! The citation has been corrected.

(11) The authors have shown that Mg+2 does not alter the LPS transport system, however, there was some significant increase in LPS expression at 200mM MgCl2. It would be interesting if the authors could also check if Mg+2 has any effect on the outer membrane protein (OMP) integrity (by checking OMP components BamA and LptD).

We appreciate this comment! We have carefully examined the membrane permeability in Figure 7. We thus didn’t perform additional experiment here to see the change of BamA and LptD. Thank you very much for your understanding.

(12) I wonder if the authors could check the effect of extracellular Mg+2 during the co-treatment of palmitic acid, linoleic acid, and balofloxacin. Will there still be the antagonistic effect or the presence of Mg+2 could change the phenotype?

We appreciate this comment! Additional experiments is performed as following:

“Furthermore, magnesium had a minimal effect on the antagonistic effect of palmitic acid, linolenic acid, and balofloxacin (Fig 4G), suggesting that this mineral functions through lipid metabolism.”

Reviewer #2 (Recommendations For The Authors):

(1) As mentioned in the Public Review, I strongly believe that the impact of this study will be more significant if magnesium-induced phenotypic drug resistance could be demonstrated in at least one other Gram-negative and one other Grampositive species, both of which should be human pathogens. The full suite of experiments would not be necessary for this suggestion; evaluation of the effect of Mg concentration in growth media on the drug resistance of other species, testing the different antibiotic types used in this study, would be sufficient.

We appreciate this comment! Additional experiments have performed to test this idea. Mg2+ has the similar effect on carbapenem-resistant Escherichia coli, carbapenem-resistant Klebsiella pneumoniae, carbapenem-resistant Pseudomonas aeruginosa and carbapenem-resistant Acinetobacter baumannii as the similar as on the Vibrio species in shown in Figure 1G. These have been described following as

“Importantly, exogenous MgCl2 also increased MICs of clinic isolates, carbapenemresistant Escherichia coli, carbapenem-resistant Klebsiella pneumoniae, carbapenemresistant Pseudomonas aeruginosa and carbapenem-resistant Acinetobacter baumannii to balofloxacin (Fig 1G).”

(2) I recommend that the Introduction section be expanded. I recommend one or two sentences introducing the two Vibrio species selected for study. I.e. why did the authors choose these two species? What is known about their phenotypic drug resistance in the literature? Why did the authors select balofloxacin for their studies, is it a common antimicrobial used vs Vibrios? As well, the end of the Introduction section ends abruptly with no transition to the present study itself. The end of the introduction should include one or two sentences introducing the main purpose of the study, its approach, and the techniques undertaken. For example, "In this study, we evaluated whether magnesium induces phenotypic resistance in Vibrio species and the molecular/genetic basis for such resistance. We used genetic approaches, GC-MS analysis of metabolite and membrane remodeling upon antibiotic exposure, membrane physiology, and extensive antimicrobial susceptibility evaluations."

We appreciate this comment! We revise the introduction by providing additional information as following:

“In Gram-negative bacteria, by contrast, zinc enhances antibiotic efficacy by potentiating carbapenem, fluoroquinolone, and β-lactam-mediated killing (Isaei et al., 2016; Zhang et al., 2014). Magnesium influences bacterial structure, cell motility, enzyme function, cell signaling, and pathogenesis (Wang et al., 2019). This mineral also modulates microbiota to harvest energy from the diet (Garcia-Legorreta et al., 2020), allowing Bacillus subtilis to cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al., 2019). However, the role of magnesium in promoting phenotypic resistance is less well understood.

Vibrios inhabit seawater, estuaries, bays, and coastal waters, regions full of metal ions such as magnesium (Kumarage et al., 2022). Magnesium is the second most dissolved element in seawater after sodium. At a salinity of 3.5% seawater, the magnesium concentration is about 54 mM (Potis, 1968), and in deep seawater, can be as high as 2,500 mM (Wang et al., 2024). Vibrio parahaemolyticus and V. alginilyticus are two representative Vibrio pathogens that infect humans and aquatic animals, resulting in illness and economic loss, respectively (Grimes, 2020). (Fluoro)quinolones such as balofloxacin are used to treat Vibrio infection, however, resistance has emerged due to overuse (Suyamud et al., 2024). Indeed, (fluoro)quinolones are one of China's two primary residual chemicals associated with aquaculture (Liu et al., 2017). Vibrio can develop quinolone resistance through mutations in the DNA gyrase gene or through plasmid-mediated mechanisms (Dutta et al., 2021). Thus, the use of V. parahaemolyticus and V. alginilyticus as bacterial representatives, and balofloxacin as a quinolone-based antibacterial representative, can help to define novel magnesiumdependent phenotypic resistance mechanisms of pathogenic Vibrio species.

The current study evaluated whether magnesium induces phenotypic resistance in Vibrio species and defined the molecular/genetic basis for this resistance. Genetic approaches, GC-MS analysis of metabolite and membrane remodeling upon antibiotic exposure, membrane physiology, and extensive antimicrobial susceptibility testing were used for the evaluations. ”

(3) The authors introduce the acronym AWST but never use it again in the paper, instead they use SWT. The authors should introduce SWT only for consistency.

We appreciate this comment! We have corrected all the “SWT” to “ASWT”

(4) Line 76 is not clear: what is meant by "some of which could influence drug efficacy" - the enzymes that utilize light metal ions are co-factors? Or the metals directly?

We appreciate this comment! The information we wanted to deliver is that light metal ions can serve as cofactors to catalyze biochemical reaction. Such chemical reaction would alter the drug efficacy, e.g. the Fe-S cluster are metallocofactor for proteins which regulates redox chemistry including antibioticinduced redox change. However, this information is not appropriate for this manuscript, so we delete this sentence.

(5) Line 90: add a reference corroborating that this chemical composition is a mimic of marine water. The NaCl concentration used in particular looks quite low.

We appreciate this comment! It was a typo error. The NaCl concentration was 210 mM as shown in Suppl. Table 1. We also provide details of the chemical composition of the marine water as following:

“Marine environments and agriculture, where antibiotics are commonly used, are rich in magnesium. To investigate whether this mineral impacts antibiotic activity, the minimal inhibitory concentration (MIC) of V. alginolyticus ATCC33787 and V. parahaemolyticus VP01, which we referred as ATCC33787 and VP01 afterwards, isolated from marine aquaculture, to balofloxacin (BLFX) in Luria-Bertani medium

(LB medium) plus 3% NaCl as LBS medium and “artificial seawater” (ASWT) medium that included the major ion species in marine water (Wilson, 1975) (LB medium plus 210 mM NaCl, 35 mM Mg2SO4, 7 mM KCl, and 7 mM CaCl2) were assessed”

(6) Line 98 and Figure 1B. M9 is indicated in the text but does not appear in the figure, the figure only shows SWT. This should be checked. Line 99: based on Figure 1C, the authors are adding MgCl2 to SWT, SWT should be mentioned in this line. Line 100: I believe this is referring to Figure 1C, which should be checked.

We appreciate this comment!

Line 98, which is now Line 118: We have corrected M9 to ASWT as following:

“However, the MIC for BLFX was higher in ASWT medium supplemented with Mg2SO4 or MgCl2 than in LB medium (Fig 1B).”

Line 99, which is now Line 133: the sentence is corrected as following:

“The MIC for BLFX increased at higher concentrations of MgCl2 in ASWT”

Line 100, which is now Line 135: we have corrected Fig 1B to Fig. 1C.

(7) Line 101: text and Figure 1D are not consistent, as Figure 1D does not show this level of precision in added MgCl2 as indicated in the text (15.6 - 62.4 mM).

We appreciate this comment! The sentence has been corrected as following: “At balofloxacin doses of 1.56, 3.125, 6.25, and 12.5 µg, the zone of inhibition decreased with increasing MgCl2 (Fig 1D)””.

(8) MgCl2 clearly induces increasing levels of BLFX resistance, and to high levels, but not for every antibiotic. For example, the level of increased resistance to blactams is low (ceftriaxone) and plateaus (ceftazidime). As well, resistance to gentamicin plateaus at a lower level than the other aminoglycosides. These observations do not take away from the conclusion that Mg induces multi-drug resistance, but since the behaviour of the MICs for these drugs is different than the other drugs, they should be mentioned. Also, Figure 1F - tetracyclines (plural) is used for vertical axis label - does this refer to the tetracycline itself or the class itself, and if the class, which one was tested?

We appreciate this comment! We revise the description as following: “Notably, magnesium had a reduced effect on ceftriaxone and gentamicin than other antibiotics.”

The tetracyclines is labeled as “Oxytetracycline” in the revised manuscript.

- The magnesium chelation experiments presented in Figure 2 are not clear. The authors should briefly mention how this was done around line 128, and what data underlies the values in Figure 2C. Figure 2B is also not clear to me at all. Similarly, how the authors measured intracellular balofloxacin and Mg2+ is not clear and should be mentioned briefly around lines 130-132.

We appreciate this comment! These have been rewritten following as “To investigate whether magnesium binds to balofloxacin, balofloxacin was preincubated with magnesium, and zone of inhibition (ZOI) analysis was conducted. Six different concentrations of balofloxacin were separately incubated with six different concentrations of MgCl2, and then spotted on filter paper so that a defined amount of balofloxacin could be used for ZOI. While lower concentrations of MgCl2, (0.78, 3.125, or 12.5 mM) did not alter the ZOI, higher concentrations, including 50 and 200 mM MgCl2, decreased the ZOI (Suppl. Fig 2A), suggesting that even high doses of magnesium had only a partial effect on balofloxacin through direct binding. For example, at 200 mM MgCl2 and 5 or 10 μg/mL balofloxacin, the balofloxacin ZOI was 53.2 and 70.3% of the ZOI at 0 mM MgCl2, suggesting that 50% of the antibiotics were still functional. Intracellular BLFX also decreased with increasing MgCl2 (Suppl. Fig 2B), while exogenous Mg2+ increased intracellular Mg2+ levels in a dose-dependent manner. For example, exogenous 50 and 200 mM MgCl2 increased intracellular Mg2+ levels to 1.21 and 1.31 mM, respectively (Suppl. Fig 2C). The relationship between TolC, an efflux pump that transports quinolones from bacterial cells, and Mg2+ was also assessed (Kobylka et al., 2020; Song et al., 2020). The expression of TolC/tolC was unaffected by Mg2+ (Suppl. Fig 2D). Magnesium is critical for LPS stability. LPS levels increased at 200 mM Mg2+ (Suppl. Fig 2E), however, the loss of waaF, lpxA, and lpxC, three key genes involved in LPS biosynthesis, did not influence balofloxacin sensitivity/resistance in the presence of Mg2+ (Suppl. Fig 2F). These findings suggest that magnesium-induced LPS biosynthesis does not contribute directly to BLFX resistance and demonstrate that Mg2+ influx is involved in balofloxacin resistance.”

- Line 135: LPS cannot be "expressed", as the authors word it here. This should be corrected. Also, the inspection of Figure 2G actually shows the levels of LPS increase with increased Mg2+. The authors should re-evaluate these results and change their description around this area of the Results.

We appreciate this comment! We have removed the whole Figure 2 to Supplementary Text and Supplementary Figure 2. We rewrite this part as following: “The relationship between TolC, an efflux pump that transports quinolones from bacterial cells, and Mg2+ was also assessed (Kobylka et al., 2020; Song et al., 2020). The expression of TolC/tolC was unaffected by Mg2+ (Suppl. Fig 2D). Magnesium is critical for LPS stability. LPS levels increased at 200 mM Mg2+ (Suppl. Fig 2E), however, the loss of waaF, lpxA, and lpxC, three key genes involved in LPS biosynthesis, did not influence balofloxacin sensitivity/resistance in the presence of Mg2+ (Suppl. Fig 2F). These findings suggest that magnesium-induced LPS biosynthesis does not contribute directly to BLFX resistance and demonstrate that Mg2+ influx is involved in balofloxacin resistance.”

- Section: MgCl2 affects bacterial metabolism. Authors switched to M9 medium - why? This contrasts with other sections using SWT and should be explained. Also, I cannot evaluate whether the statistical analysis of the data here was performed correctly and was appropriate for this type of experiment. I advise the authors to move the details in lines 166-169 to the Materials and Methods and replace this section instead with a more accessible description of the statistical analysis that a non-expert would be able to appreciate. Furthermore, analysis of Figure 3A indicates that the levels of asparagine, 4-hydroxybutyric acid, uracil, cystathionine, fumaric acid, and aminoethanol have significantly changed at high MgCl2, but these are not mentioned in the text. I suggest the authors mention these if they are relevant to the 12 enriched pathways, especially the biosynthesis of fatty acids.

We appreciate this comment!

We indicate the reason we use M9 medium as following:

“To better understand how magnesium affects bacterial metabolism” for explaining why the M9 medium was used.”

The information lines 166-169 indicated has been removed to M &M.

We have carefully examined the abundance of the metabolites and the enriched pathway. Among the listed metabolites, only fumarate is within the enriched pathways. We mention this point in our revised manuscript as following:

“The increase in fatty acid biosynthesis could be partially explained by an imbalanced pyruvate cycle/TCA cycle, in which fumarate levels increased at higher Mg2+ while succinate levels increased at lower Mg2+ (Suppl. Fig 5B). These findings indicated that glycolysis fluxes into fatty acid biosynthesis rather than the pyruvate cycle/TCA cycle. The relevance of fatty acids and BLFX was demonstrated by the observation that exogenous palmitic acid increased bacterial resistance to balofloxacin (Fig 2F). These results suggest that fatty acid metabolism may be critical to magnesium-based phenotypic resistance.”

- Line 211 appears to refer to Figure 4F and should be checked. Similarly in line 216 - appears this should be Figure 4H, and line 218 should be Figure 4H. Line 226: add a reference to Fig 4I (after arcA was decreased). Line 227: what are genes N646_1004 and N646_1885? Based on Fig 4J these are crp - authors should add to line 227. Line 228 appears to refer to Figure 4J, not Figure 4I. Line 229 - should be Figure 4K, not Figure 4I. Line 231 - should be 4L, not 4K. Line 239 - should be 4M.

We appreciate this comment! The text and figure is now matched.

- Line 312: the descriptions of "11 lipids, 32 lipids, and 53", and then "26 lipids, 52 lipids, and 107 lipids" are not clear at all and should be corrected.

We appreciate this comment! The sentence is revised as following:

“The abundance of 11, 32, and 53 lipids was increased in 3.125, 50, and 200 mM MgCl2-treated bacteria, respectively, while the abundance of 26, 52, and 107 lipids was decreased in 3.125, 50, and 200 mM MgCl2-treated bacteria, respectively (Suppl. Fig 7C)”

- Line 340. What is the assay the authors are using to measure the levels of the PGS and PSS enzymes? This is not mentioned or clear in this part of the Results.

We appreciate this comment! We provide the information in the manuscript as following:

“Levels of PGS and PSS were quantified by ELISA kits according to manufacture’s instruction (Shanghai Fusheng Industrial Co., Ltd., China)”

- Line 372: What is the assay for measuring membrane depolarization? This is not mentioned and I suggest it should be. Line 374: Figure 7B does not show time dependence, only dose dependence, this should be corrected, it is assumed the authors are referring to Fig 7C for the time dependence data.

We appreciate this comment! We provide the information in the result as following:

“The voltage-sensitive dye, DiBAC4(3) showed that 12.5–200 mM MgCl2 promoted membrane depolarization in a dose-dependent manner (Fig 6A)”

We also explain how DiBAC4(3) can be used to measure membrane depolarization in the Materials and Methods section as following:

“DiBAC4(3) is a s voltage-sensitive probe that penetrates depolarized cells, binding intracellular proteins or membranes exhibiting enhanced fluorescence and red spectral shift.”

To make it clear the specific figure, we revise the sentence as following:

“Meanwhile, MgCl2 had a dose-dependent (Fig 6B) and time-dependent (Fig 6C) effect on proton motive force (PMF).”

- Line 384: mention how FM5-95 measures membrane permeability. The authors should also clarify how this reagent is used to measure membrane fluidity, and it is not clear if the data for this is presented in Figure 7 - please clarify. Regarding SYTO9 dye experiment: the authors should briefly explain the experimental design - how SYTO9 dye operates and why FACS was chosen. What is labeled with FITC?

We appreciate this comment! We clarify the reason we use FM5-95 in the Methods and Materials section as following:

“Measurement of fluidity by fluorescence microscopy

Measurement of membrane fluidity is performed as previously described (Wen et al., 2022). Briefly, ATCC33787 were cultured in medium with indicated concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacteria cells of each sample were diluted to 1 mL and 10 μL (10 mg/mL) FM5-95 (Thermo Fisher

Scientific, USA) was added. FM5-95 is a lipophilic styryl dye that insert into the outer leaflet of bacterial membrane and become fluorescence. This dye preferentially bind to the microdomains with high membrane fluidity(Wen et al., 2022). After incubated for 20 min at 30 ℃ at vibration without light, the sample was centrifuged for 10 min at 12,000 rpm. The pellets were resuspended with 20 μL of 3% NaCI. Aliquot of 2 μL sample was dropped on the agarose slide, and take photos under the inverted fluorescence microscope.”

This data is presented as micrographs in Fig. 6D, which shows the decreased FM5-95 staining with increasing concentrations of MgCl2. We make this description clear with the following revision:

“FM5-95 staining decreased with increasing concentrations of Mg2+, and no staining was observed in the presence of 200 mM Mg2+ (Fig 6D).”

We explain the reason why we use SYTO9 as following:

“SYTO9, a green fluorescent dye that binds to nucleic acid, enters and stains bacteria cells when there is an increase in membrane permeability (Lehtinen et al., 2004; McGoverin et al., 2020). Staining decreased with increasing MgCl2, indicating that bacterial membrane permeability declined in an Mg2+ dose-dependent manner (Fig 6E).”

We didn’t use FACS in this study, while we only analyze the fluorescence distribution with the equipment. To make it clear, we revise the sentence as following:

“After incubated for 15 min at 30 ℃ at vibration without light, the mixtures were filtered and measured by flow cytometry (BD FACSCalibur, USA).”

- Lines 391-397. The statement that palmitic acid shifts the peaks in Figure 7F is not supported by the data. There is essential no change in the major peak position within each MgCl2 concentration set with increasing palmitic acid. For the linolenic acid data, it is clear that linolenic acid increases permeability only at 50 mM MgCl2-this should be mentioned in the text.

We appreciate this comment! We revise the sentence as following:

“Exogenous palmitic acid also shifted the fluorescence signal peaks to the left in an MgCl2-dependent manner while palmitic acid only slightly shifted the peaks (Fig 6F). In contrast, exogenous linolenic acid shifted the peak to the right in a dose-dependent manner at 50 mM MgCl2 (Fig 6G).”

- Line 404-405 - as mentioned earlier, the assay for the update of BLFX should be mentioned (if it is done so earlier in the text, then it does not need to be here).

We appreciate this comment! It has been mentioned in the introduction.

- Discussion: CpxA/R-OmprF pathway is mentioned here for the first time. Is this one of the pathways modified by MgCl2 as determined during the course of the study? If so, this should be reworded to mention that. If not, the relevance of this particular pathway as it relates to light metals and phenotypic resistance should be discussed.

We appreciate this comment! Since it is not relevant to the discussion of Mg2+ and fatty acid biosynthesis, we delete this sentence in the revised manuscript.

-The following grammatical errors should be corrected:

-line 55 change to: "genetic mutations; instead, this type of resistance is transient, and bacteria resume normal growth"

-line 57: change to "resistance types are biofilm"

-line 61: change to "states that significantly"

-line 63: change to "resistance share the common feature in they retard or even cease in the presence"

-line 65: change to "resistance that allow bacteria to proliferate"

-line 81: change "But whether" to "Whether"

-line 178: change to "may be critical to the Mg-based phenotypic resistance"

-line 86: change to "Marine environments and agriculture are rich in magnesium, where..."

-line 93: change in to vs

-line 154: insert space after metabolism

-line 158: change 'identified" to "focused on the levels of"

-line 160: change "The levels of forty-one metabolites"

-line 198: change shared to share

-line 310: increased is duplicated, delete one

-line 451: add "the" before ratio

-line 453: gram should be capitalized

-line 462: "the regulation" should be reworded to "More importantly, the effect of exogenous MgCl targets the..."

-line 469: add dash between Mg2+ and limited

-line 478: change "the crucial" to "a crucial"

-there are numerous locations in the manuscript where the word "magnetism" is used when clearly the word is supposed to be magnesium - this should be corrected

We appreciate this comment! These have been corrected or revised.

Editors comments:

Page 2 line 27; Page 25 line number 426; page 27 line number 481: In the abstract and discussion, only Vibrio alginolyticus was mentioned, even though two Vibrio species were used in the study. It would be helpful to understand the rationale behind the focus on this particular species.

We appreciate this comment! We have revised the introduction to provide additional information as following:

“Vibrios inhabit seawater, estuaries, bays, and coastal waters, regions full of metal ions such as magnesium (Kumarage et al., 2022). Magnesium is the second most dissolved element in seawater after sodium. At a salinity of 3.5% seawater, the magnesium concentration is about 54 mM (Potis, 1968), and in deep seawater, can be as high as 2,500 mM (Wang et al., 2024). Vibrio parahaemolyticus and V. alginilyticus are two representative Vibrio pathogens that infect humans and aquatic animals, resulting in illness and economic loss, respectively (Grimes, 2020). (Fluoro)quinolones such as balofloxacin are used to treat Vibrio infection, however, resistance has emerged due to overuse (Suyamud et al., 2024). Indeed, (fluoro)quinolones are one of China's two primary residual chemicals associated with aquaculture (Liu et al., 2017). Vibrio can develop quinolone resistance through mutations in the DNA gyrase gene or through plasmid-mediated mechanisms (Dutta et al., 2021). Thus, the use of V. parahaemolyticus and V. alginilyticus as bacterial representatives, and balofloxacin as a quinolone-based antibacterial representative, can help to define novel magnesium-dependent phenotypic resistance mechanisms of pathogenic Vibrio species.”

On Page 2, line 34: The abstract contains some undefined abbreviations, such as 'PE' and 'PG', which should be explained.

We appreciate this comment! We explain the PE and PG in the revised abstract as following:

“phosphatidylethanolamine (PE) biosynthesis is reduced and phosphatidylglycerol (PG)”

On Page 2, line 31-32: For the statement "Exogenous supplementation of fatty acids confirm the role of fatty acids in antibiotic resistance…" it would be beneficial to specify whether the fatty acids were saturated or unsaturated.

Response, We appreciate this comment! We revise the sentence as following:

“Exogenous supplementation of unsaturated and saturated fatty acids increased and decreased bacterial susceptibility to antibiotics, respectively, confirming the role of fatty acids in antibiotic resistance.”

The potential effects of the specific ions (SO4 and Cl2) present in the Mg2SO4 and MgCl2 compounds used in the study were not discussed. It would be useful to understand if these ions had any influence on the observed outcomes.

We appreciate this comment! We revise the sentence as following:

“However, the MIC for BLFX was higher in ASWT medium supplemented with Mg2SO4 or MgCl2 than in LB medium (Fig 1B). And Mg2SO4 or MgCl2 had no

difference on MIC, suggesting it is Mg2+ not other ions contribute to the MIC change.”

On Page 8, line 141: The heading of Figure 2, "Mg2+ elevates intracellular Mg2+," seems redundant and could be revised for clarity or modified.

We appreciate this comment! Figure 2 is now moved to supplementary figure as Suppl. Fig 2. The title is revised as following:

“Figure 2. Mg2+ decreases balofloxacin uptake.”

On Page 4, line 91: some terms/abbreviations, such as 'LB' and 'M9,' require expansion or definition to ensure the reader's understanding.

We appreciate this comment! We include the expansion for LB and M9 in the revised manuscript as following:

“Luria-Bertani medium (LB medium)” and “M9 minimal medium (M9 medium)”

Page 4, line 92: The real seawater composition used in the experiments should be supported by a reference.

We appreciate this comment! We provide the reference in the revised manuscript as following:

““artificial seawater” (ASWT) medium that included the major ion species in marine water (Wilson, 1975) (LB medium plus 210 mM NaCl, 35 mM Mg2SO4, 7 mM KCl, and 7 mM CaCl2)”

Page 4 line, number 93: the he full names of the bacterial strains (e.g., ATCC33787 and VP01) should be provided instead of just the strain numbers.

We appreciate this comment! We revised the sentence as following:

“To investigate whether this mineral impacts antibiotic activity, the minimal inhibitory concentration (MIC) of V. alginolyticus ATCC33787 and V. parahaemolyticus VP01, which we referred as ATCC33787 and VP01 afterwards,”

Finally, there appears to be a potential contradiction between the statements on page 12, lines 211-212 and 214-216, regarding the effects of Mg2+ on the synthesis of unsaturated fatty acids. Further explanation may be needed to reconcile these seemingly contradictory points.

We appreciate this comment! For line 221-226, which was previously line 211-212, is about the gene expression for fatty acid biosynthesis. While, Line 228 and 233, which was previously line 214-216 is about the gene expression for fatty acid degradation. We agree that the previous description is a little bit confuse. We revise the sentence to emphasize that we focus on fatty acid degradation so that the readers can tell them apart.

In the text, we revised it as following:

“In addition, we also quantified gene expression during fatty acid degradation to determine whether Mg2+ affects this process” In the figure legend, we also indicate that

“H. qRT-PCR for the expression of genes encoding fatty acid degradation in the absence or presence of the indicated concentrations of MgCl2”

Version 2

Abstract

Abstract

Non-inheritable antibiotic or phenotypic resistance ensures bacterial survival during antibiotic treatment. However, exogenous factors promoting phenotypic resistance are poorly defined. Here, we demonstrate that Vibrio alginolyticus are recalcitrant to killing by a broad spectrum of antibiotics under high magnesium. Functional metabolomics demonstrated that magnesium modulates fatty acid biosynthesis by increasing saturated fatty acid biosynthesis while decreasing unsaturated fatty acid production. Exogenous supplementation of unsaturated and saturated fatty acids increased and decreased bacterial susceptibility to antibiotics, respectively, confirming the role of fatty acids in antibiotic resistance. Functional lipidomics revealed that glycerophospholipid metabolism is the major metabolic pathway remodeled by magnesium, where phosphatidylethanolamine (PE) biosynthesis is reduced and phosphatidylglycerol (PG) production is increased. This process alters membrane composition, increasing membrane polarization, and decreasing permeability and fluidity, thereby reducing antibiotic uptake by V. alginolyticus. These findings suggest the presence of a previously unrecognized metabolic mechanism by which bacteria escape antibiotic killing through the use of an environmental factor.

Introduction

Non-inheritable antibiotic or phenotypic resistance represents a serious challenge for treating bacterial infections. Phenotypic resistance does not involve genetic mutations Phenotypic resistance does not involve genetic mutations and is transient, allowing bacteria to resume normal growth. Biofilm and bacterial persisters are two phenotypic resistance types that have been extensively studied (Brandis et al., 2023; Corona & Martinez, 2013). Biofilms have complex structures, containing elements that impede antibiotic diffusion, sequestering and inhibiting their activity (Ciofu et al., 2022). Biofilm-forming bacteria and persisters also have distinct metabolic states that significantly reduce their antibiotic susceptibility (Yan & Bassler, 2019). These two types of phenotypic resistance share the common feature in their retarded or even cease of growth in the presence of antibiotics (Corona & Martinez, 2013). However, specific factors that promote phenotypic resistance and allow bacteria to proliferate in the presence of antibiotics remain poorly defined.

Metal ions have a diverse impact on the chemical, physical, and physiological processes of antibiotic resistance (Booth et al, 2011; Lu et al, 2020; Poole, 2017). This includes genetic elements that confer resistance to metals and antibiotics (Poole, 2017) and metal cations that directly hinder (or enhance) the activity of specific antibiotic drugs (Zhang et al., 2014). The metabolic environment can also impact the sensitivity of bacteria to antibiotics (Jiang et al., 2023; Lee & Collins, 2012; Peng et al., 2015; Zhang et al., 2020; Zhao et al., 2021). Light metal ions, such as magnesium, sodium, and potassium, can behave as cofactors for different enzymes (Du et al., 2016) and influence drug efficacy. Heavy metal ions, including Cu2+ and Zn2+, confer resistance to antibiotics (Yazdankhah et al., 2014; Zhang et al., 2018). Recent reports suggest that sodium negatively regulates redox states to promote the antibiotic resistance of Vibrio alginolyticus (Yang et al., 2018), while actively growing Bacillus subtilis cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al, 2019). In Gram-negative bacteria, by contrast, zinc enhances antibiotic efficacy by potentiating carbapenem, fluoroquinolone, and β-lactam-mediated killing (Isaei et al., 2016; Zhang et al., 2014). Magnesium influences bacterial structure, cell motility, enzyme function, cell signaling, and pathogenesis (Wang et al., 2019). This mineral also modulates microbiota to harvest energy from the diet (Garcia-Legorreta et al., 2020), allowing Bacillus subtilis to cope with ribosome-targeting antibiotics by modulating ion flux (Lee et al., 2019). However, the role of magnesium in promoting phenotypic resistance is less well understood.

Vibrios inhabit seawater, estuaries, bays, and coastal waters, regions full of metal ions such as magnesium (Kumarage et al., 2022). Magnesium is the second most dissolved element in seawater after sodium. At a salinity of 3.5% seawater, the magnesium concentration is about 54 mM (Potis, 1968), and in deep seawater, can be as high as 2,500 mM (Wang et al., 2024). Vibrio parahaemolyticus and V. alginilyticus are two representative Vibrio pathogens that infect humans and aquatic animals, resulting in illness and economic loss, respectively (Grimes, 2020). (Fluoro)quinolones such as balofloxacin are used to treat Vibrio infection, however, resistance has emerged due to overuse (Suyamud et al., 2024). Indeed, (fluoro)quinolones are one of China’s two primary residual chemicals associated with aquaculture (Liu et al., 2017). Vibrio can develop quinolone resistance through mutations in the DNA gyrase gene or through plasmid-mediated mechanisms (Dutta et al., 2021). Thus, the use of V. parahaemolyticus and V. alginilyticus as bacterial representatives, and balofloxacin as a quinolone-based antibacterial representative, can help to define novel magnesium-dependent phenotypic resistance mechanisms of pathogenic Vibrio species.

The current study evaluated whether magnesium induces phenotypic resistance in Vibrio species and defined the molecular/genetic basis for this resistance. Genetic approaches, GC-MS analysis of metabolite and membrane remodeling upon antibiotic exposure, membrane physiology, and extensive antimicrobial susceptibility testing were used for the evaluations.

Results

Mg2+ promotes phenotypic resistance to antibiotics

Marine environments and agriculture, where antibiotics are commonly used, are rich in magnesium. To investigate whether this mineral impacts antibiotic activity, the minimal inhibitory concentration (MIC) of V. alginolyticus ATCC33787 and V. parahaemolyticus VP01, which we referred as ATCC33787 and VP01 afterwards, isolated from marine aquaculture, to balofloxacin (BLFX) in Luria-Bertani medium (LB medium) plus 3% NaCl as LBS medium and “artificial seawater” (ASWT) medium that included the major ion species in marine water (Wilson, 1975) (LB medium plus 210 mM NaCl, 35 mM Mg2SO4, 7 mM KCl, and 7 mM CaCl2) were assessed (Suppl. Table 1). The MICs of ATCC33787 to BLFX were 54 and 1 μg/mL in ASWT and LBS medium, respectively. Similarly, the MICs to VP01 were 25 and 3 μg/mL in ASWT and LBS medium, respectively (Fig 1A). The role of exogenous NaCl in antibiotic resistance was investigated previously (Yang et al., 2018). The current study found that the MIC for BLFX was the same in LB and M9 minimal medium (M9 medium) plus 7 mM KCl or 7 mM CaCl2 (Suppl. Fig 1). However, the MIC for BLFX was higher in ASWT medium supplemented with Mg2SO4 or MgCl2 than in LB medium (Fig 1B). And Mg2SO4 or MgCl2 had no difference on MIC, suggesting it is Mg2+ not other ions contribute to the MIC change. The MIC for BLFX increased at higher concentrations of MgCl2 in ASWT medium. Specifically, adding 50 mM or 200 mM MgCl2 increased the MIC for BLFX by 200- or 1600-fold, respectively (Fig 1C). At balofloxacin doses of 1.56, 3.125, 6.25, and 12.5 µg, the zone of inhibition decreased with increasing MgCl2 (Fig 1D). Exogenous MgCl2 also increased the MICs for other quinolone (e.g. nalidixic acid, levofloxacin, ciprofloxacin, ofloxacin, and moxifloxacin) (Fig 1E) and non-quinolone antibiotics including antibacterial peptides (colistin), macrolides (roxithromycin), tetracyclines (oxytetraycline), β-lactams (ceftriaxone, ceftazidime), and aminoglycosides (amikacin, kanamycin, and gentamicin) (Fig 1F). Notably, magnesium had a reduced effect on ceftriaxone and gentamicin than other antibiotics. Importantly, exogenous MgCl2 also increased MICs of clinic isolates, carbapenem-resistant Escherichia coli, carbapenem-resistant Klebsiella pneumoniae, carbapenem-resistant Pseudomonas aeruginosa and carbapenem-resistant Acinetobacter baumannii to balofloxacin (Fig 1G). These findings indicate that Mg2+ promotes phenotypic resistance.

Magnesium promotes bacterial resistance to antibiotics.

A. MIC of ATCC33787 and V. parahaemolyticus to BLFX in ASWT or LB medium as determined by the microtitre-dilution method. B. MIC of ATCC33787 to BLFX in ASWT with additional MgSO4 or MgCl2 as determined by the microtitre-dilution-method. C. MIC of ATCC33787 to BLFX in ASWT at the indicated concentrations of MgCl2 as determined by the microtitre-dilution method. D. MIC of ATCC33787 to BLFX at the indicated concentrations of BLFX and MgCl2 as determined by the Oxford cup test. The numbers 1, 2, 3, 4, 5, 6, and 7 represent 0, 0.39, 0.78, 1.56, 3.125, 6.25, and 12.5 µg BLFX, respectively. E. MIC of ATCC33787 to other quinolones in ASWT at the indicated concentrations of MgCl2 as determined by the microtitre-dilution method. F. MIC of ATCC 33787 to other classes of antibiotics in ASWT at the indicated concentrations of MgCl2 as determined by the microtitre-dilution method. G. MIC of carbapenem-resistant Escherichia coli, carbapenem-resistant Klebsiella pneumoniae, carbapenem-resistant Pseudomonas aeruginosa, and carbapenem-resistant Acinetobacter baumannii isolates to BFLX at the indicated concentrations of MgCl2.

MgCl2 affects bacterial metabolism

To investigate how magnesium regulates phenotypic resistance, other effects of magnesium, including its ability to quench balofloxacin activity and induce efflux pump expression and LPS biosynthesis, were excluded (Suppl. Text; Suppl. Fig 2). While Mg2+ is a co-factor for enzymes (Garfinkel and Garfinkel, 1985), whether this mineral can promote antibiotic resistance independent of this function remains unknown.

To better understand how magnesium affects bacterial metabolism, V. alginolyticus was cultured in M9 medium in the presence of various amounts of MgCl2 (0, 0.78, 3.125, 12.5, 50, or 200 mM) and the levels of 54 metabolites were assessed by GC-MS. Five biological and two technical replicates were evaluated for each treatment (Suppl. Fig 3). The levels of 41 metabolites were differential (p <0.05) and are presented as Z-values (Fig 2A; Suppl. Fig 4). Orthogonal partial least square discriminant analysis (OPLS-DA) was conducted that separated the six treatments into three groups: Group 1 included 0 mM, 0.78 mM, and 3.125 mM MgCl2; Group 2 was a singlet of 12.5 mM; Group 3 included 50 mM and 200 mM MgCl2. Component t[1] distinguishes Group 1 from Groups 2 and 3; component t[2] distinguishes Group 3 from Groups 1 and 2 (Fig 2B). Discriminating variables are shown as S-plot, where cut-off values are ≥ 0.05 absolute value of covariance p and 0.5 for correlation p(corr). Six biomarkers/metabolites were selected from component t[1] and p[1]. More specifically, the abundance of cadaverine, urea, palmitic acid, aminoethanol, and fumaric acid were increased, but pyroglutamic acid and glutamic acid were decreased (Fig 2C; Suppl. Fig 5A). Pathway enrichment analysis suggests that twelve pathways are involved. Notably, one of the pathways is biosynthesis of unsaturated fatty acids (terminology in the software. Actually, it is biosynthesis of fatty acids) (Fig 2D), where the abundance of palmitic acid and stearic acid were increased in an Mg2+ dose-dependent manner (Fig 2E; Suppl. Fig 6). Moreover, palmitic acid was a crucial biomarker (Suppl. Fig 6). The increase in fatty acid biosynthesis could be partially explained by an imbalanced pyruvate cycle/TCA cycle, in which fumarate levels increased at higher Mg2+ while succinate levels increased at lower Mg2+ (Suppl. Fig 5B). These findings indicated that glycolysis fluxes into fatty acid biosynthesis rather than the pyruvate cycle/TCA cycle. The relevance of fatty acids and BLFX was demonstrated by the observation that exogenous palmitic acid increased bacterial resistance to balofloxacin (Fig 2F). These results suggest that fatty acid metabolism may be critical to magnesium-based phenotypic resistance.

Mg2+-induced metabolomic change.

A. Differential metabolomes in the absence or presence of the indicated concentrations of MgCl2. The yellow and blue colors indicate an increase or decrease in metabolite levels relative to the median metabolite level, respectively (see color scale). Euclidean distance calculations were used to generate a heatmap that shows clustering of the biological and technical replicates of each treatment. B. OPLS-DA analysis of different MgCl2-induced metabolome concentrations. Each dot represents a technical replicate of samples in the plot. C. S-plot generated from OPLS-DA. Predictive component p [1] and correlation p(corr) [1] differentiate 0, 0.78, 3.125 mM MgCl2 from 12.5, 50, 200 mM MgCl2. Predictive component p[2] and correlation p(corr)[2] separate 0, 0.78, 50, 200 mM MgCl2 from 3.125, 12.5 mM MgCl2. The triangle represents metabolites in which candidate biomarkers are marked. D. Enriched pathways by differential abundances of metabolites. E. Areas of the peaks of palmitic acid and stearic acid generated by GC-MS analysis. F. Synergy analysis for BFLX with palmitic acid for V. alginolyticus. Synergy was performed by comparing the dose needed for 50% inhibition of the synergistic agents (white) and non-synergistic (i.e., additive) agents (purple).

Mg2+ regulates fatty acid biosynthesis

Acetyl-CoA carboxylase (ACC) catalyzes the conversion of acetyl-CoA to malonyl-CoA, the first step of fatty acid synthesis. Fatty acid biosynthesis, which includes saturated and unsaturated forms, shares common biosynthetic pathways with acetyl-CoA to enoyl-ACP, and is sequentially mediated by fabD or fabH, fabB/F, fabG, and fabA/Z. The enzyme, enoyl-ACP, produces unsaturated fatty acids or is metabolized to acyl-ACP via fabV to produce saturated fatty acids. Unsaturated fatty acids produce saturated fatty acids via tesA/B and yciF (Fig 3A). qRT-PCR was used to quantify the expression of involved genes in bacteria treated with different concentrations of MgCl2. The expression of 18 genes increased following treatment with 50 and/or 200 mM MgCl2 (Fig 3B). The expression of tesA, tesB, and yciA, which convert unsaturated to saturated fatty acids, increased at 50 or/and 200 mM MgCl2 (Fig 3C). FabA and FabF play key roles in unsaturated fatty acid and fatty acid biosynthesis, respectively (Feng & Cronan, 2009; Lee et al., 2013). Western blot analysis showed that FabA levels declined and FabF levels increased at higher MgCl2 concentrations (Fig 3D). Oxazole-2-amine and triclosan inhibit the synthesis of fatty acids and decrease bacterial survival in the presence of balofloxacin (Fig 3E and 3F). These results indicate that Mg2+ inhibits the synthesis of unsaturated fatty acids and promotes the synthesis of saturated fatty acids.

Mg2+ promotes fatty acid biosynthesis.

A. Diagram of fatty acid biosynthesis. B. qRT-PCR for expression of fatty acid biosynthesis genes in the absence or presence of indicated concentrations of MgCl2. C. qRT-PCR for expression of genes involved in converting unsaturated fatty acids to saturated fatty acids in the absence or presence of indicated MgCl2 levels. D. Western blot for the abundance of proteins responsible for converting unsaturated fatty acids to saturated fatty acids in the absence or presence of indicated concentrations of MgCl2. E. Diagram of saturated and unsaturated acid biosynthesis. F. Synergy analysis of balofloxacin with triclosan + oxazole-2-amine for ATCC33787. The percentage of bacterial survival was quantified in the presence of indicated MgCl2 levels and/or balofloxacin plus triclosan and oxazole-2-amine and used to construct the isobologram. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition of synergistic agents (blue) and non-synergistic (i.e., additive) agents (red). G. Diagram of fatty acid degradation. H. qRT-PCR for the expression of genes encoding fatty acid degradation in the absence or presence of the indicated concentrations of MgCl2. I. Western blot for the abundance of FadL in the absence or presence of the indicated concentrations of MgCl2. J. qRT-PCR for the expression of fatty acid biosynthesis regulatory genes in the absence or presence of the indicated concentrations of MgCl2. K. Western blot for the abundance of FadR in the absence or presence of the indicated concentrations of MgCl2. L. Synergy analysis for MgCl2 with BLFX for ATCC33787 (WT), ΔfadR, and ΔarcA. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition of the synergistic agents (blue) and non-synergistic (i.e., additive) agents (red). M, N, and O qRT-PCR for expression of genes involved in fatty acid biosynthesis (M) and degradation (N and O) in ATCC33787, ΔfadR or/and ΔarcA (O) in the presence or absence of 200 mM MgCl2. Whole cell lysates resolved by SDS-PAGE gel was stained with Coomassie brilliant blue as loading control (D), (I) and (K).

In addition, we also quantified gene expression during fatty acid degradation to determine whether Mg2+ affects this process (Fig 3G and 3H). Interestingly, the expression of genes involved in unsaturated fatty acid degradation decreased in an MgCl2 dose-dependent manner (Fig 3H). Western blot analysis confirmed that FadL levels decreased with increasing MgCl2 (Fig 3I). FadH degrades unsaturated fatty acids and influences the ratio of unsaturated to saturated fatty acids. Together, these results suggest that MgCl2 inhibits fatty acid degradation.

FabR, FadR, ArcA, and cAMP/CRP are transcriptional factors involved in regulating fatty acid biosynthesis. FabR inhibits unsaturated fatty acid biosynthesis, FadR promotes fatty acid biosynthesis and inhibits degradation, and ArcA and cAMP/CRP inhibit and promote fatty acid degradation, respectively (Feng and Cronan, 2010; Fujita et al., 2007). The expression of fabR and arcA increased, while the expression of fadR decreased. The expression of N646_1885, which encodes CRP, increased only in the presence of 200 mM MgCl2 (Fig 3J). Western blot analysis confirmed that FadR levels declined with increasing MgCl2 concentration (Fig 3K). Viability of ΔfadR and ΔarcA was reduced in the presence of MgCl2, balofloxacin, or both in a dose-dependent manner (Fig 3L). Thus, FadR and ArcA may play roles in bacterial resistance to balofloxacin, presumably by a mechanism dependent on the abundance of saturated/unsaturated fatty acids.

To investigate whether Mg2+ functions through FadR and ArcA, expression of eight fatty acid biosynthetic genes (accA, accC, fabD, fabH, fabR, fabG, fabV, and yciA) and five fatty acid degradation genes (fadL, fadD, fadB, fadA, and fadH) were quantified in wildtype, ΔfadR or/and ΔarcA bacterial strains. As expected, fatty acid biosynthetic gene expression was reduced in ΔfadR cells (Fig 3M). While higher expression of fadD, fadB, and fadA (elevated fadA only for ΔfadR without 200 mM MgCl2) was detected in ΔfadR or ΔacrA strains, the expression of these genes was negatively regulated by 200 mM MgCl2 (Fig 3N and 3O). These results indicate that Mg2+ promotes fatty acid biosynthesis by positively regulating fadR, and inhibits the degradation of fatty acids through fadR and arcA.

MgCl2 affects 16-carbon and 18-carbon fatty acid metabolism

To investigate the effect of Mg2+ on the biosynthesis of saturated and unsaturated fatty acids, LC-MS was used to quantify the levels of 16-carbon and 18-carbon fatty acids, the main precursors of lipid biosynthesis (Zhang and Rock, 2008b). The abundance of the saturated fatty acid, palmitic acid (C16:0), was higher while the abundance of five unsaturated fatty acids, palmitoleic acid (C16:1), linoelaidic aid (C18:2), linoleic acid (C18:2), alpha-linoleic acid (C18:3), and stearidonic acid (C18:4) was reduced with increasing MgCl2 concentration (Fig 4A). The levels of stearic acid (C16:0) and vaccenic acid (C18:1) remained unaffected. Total saturated fatty acid levels increased and total unsaturated fatty acid levels decreased with increasing Mg2+ (Fig 4B). These results suggest that Mg2+ upregulates saturated fatty acid biosynthesis but downregulates unsaturated fatty acid biosynthesis.

LC-MS targeted 16-carbon and 18-carbon fatty acids and the role of palmitic acids and linolenic acids in BLFX resistance.

A. Scatter plots of 16-carbon and 18-carbon fatty acids, detected by LC-MS. The area indicates the area of the peak of the metabolite in total ion chromatography using GC-MS. B. Scatter plots of total saturated fatty acids and unsaturated fatty acids with 16-carbon and 18-carbon (A). C. Synergy analysis for BLFX with linolenic acid for ATCC 33787. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition of synergistic agents (while) and non-synergistic (i.e., additive) agents (purple). D. LC-MS for the abundance of intracellular linolenic acid and palmitic acid of ATCC33787 in synergy with the indicated exogenous linolenic acid and palmitic acid. E. Synergy analysis of linolenic acid and palmitic acid in BLFX-mediated killing to ATCC33787. Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition of synergistic agents (blue) and non-synergistic (i.e., additive) agents (red). F. Bliss analysis (E). G. Percent survival of ATCC33787 in the presence of linolenic acid, palmitic acid, and BLFX with or without 50 mM MgCl2.

Direct exposure to palmitic acid also inhibited balofloxacin-mediated killing (Fig 2F) while linolenic acid promoted balofloxacin-mediated killing in a dose-dependent manner (Fig 4C). Increasing exogenous palmitic acid and linolenic acid increased intracellular palmitic acid and linolenic acid levels, respectively (Fig 4D). When cells were co-treated with palmitic acid, linolenic acid, and balofloxacin, the two fatty acids appeared to antagonize each other, as demonstrated by the Bliss model (Fig 4E and 4F). Furthermore, magnesium had a minimal effect on the antagonistic effect of palmitic acid, linolenic acid, and balofloxacin (Fig 4G), suggesting that this mineral functions through lipid metabolism. These results indicate that saturated and unsaturated fatty acids have an antagonistic effect on balofloxacin resistance.

Mg2+ promotes phospholipid biosynthesis

Fatty acids are biosynthetic precursors of lysophosphatidic acid (LPA) and phosphatidic acid (PA), two key cell membrane components (Zhang and Rock, 2008a). Thus, LC-MS was used to assess the effect of Mg2+ on membrane lipid composition (Suppl. Fig 7A). Higher Mg2+ levels increased the percentage of lipids from 61 to 67%, and saturated fatty acids from 24 to 26%, but decreased the percentage of unsaturated fatty acids from 15 to 8% (Suppl. Fig 7B). The abundance of 11, 32, and 53 lipids was increased in 3.125, 50, and 200 mM MgCl2-treated bacteria, respectively, while the abundance of 26, 52, and 107 lipids was decreased in 3.125, 50, and 200 mM MgCl2-treated bacteria, respectively (Suppl. Fig 7C). Saturated fatty acids and lipids increased and unsaturated fatty acids decreased in an Mg2+ dose-dependent manner (Suppl. Fig 7D). A total of 52 lipids were quantified, including 17 high-abundance lipids (Suppl. Fig 7E). Phosphatidylethanolamine (PE) and phosphatidylglycerol (PG) had the first and second highest abundance (Suppl. Fig 7E) and PE levels declined while PG levels increased with increasing Mg2+ (Fig 5A and 5B). A similar effect was observed for the unsaturated derivatives of PE and PG (Fig 5C). Principal component analysis (PCA) showed that component t [1] differentiated 0 and 3.125 mM Mg2+ from 50 and 200 mM Mg2+, while component t [2] separated 50 mM Mg2+ from 0 and 200 mM Mg2+ and variants in 3.125 mM of Mg2+ treatment (Fig 5D). S-plot analysis showed that behenic acid, PG[16:1(9Z)/16:1(9Z)], PG[18:2 (9Z, 12Z)/16:0], PA [14:0/22:6 (4Z, 7Z, 10Z, 13Z, 16Z, 19Z)], and lysoPE(16:0/0:0) were upregulated while linoelaidic acid, palmitoleic acid, PE-NMe(24:0/24:0), PE[16:0/14:1(9Z)], and lysoPA(i-12:0/0:0) were downregulated (Fig 5E). Of these, the abundance of PG[16:1(9Z)/16:1(9Z)], PG[18:2(9Z,12Z)/16:0], and behenic acid increased, while the levels of linoelaidic acid, palmitoleic acid, PE[16:0/14:1(9Z)], and lysoPA decreased with increasing Mg2+ levels (Fig 5F). Thus, altered phospholipid abundance may play a role in the effect of Mg2+ on antibiotic resistance.

Effect of Mg2+ on phospholipid biosynthesis.

A. Heatmap showing changes in differential lipid levels at the indicated concentration of MgCl2. B. Abundance of ATCC33787 phosphatidylglycerol (PG) and phosphatidylethanolamine (PE) at the indicated concentrations of MgCl2. C. Scatter plots of PG and PE at different saturation levels in the presence of the indicated MgCl2 levels. D. PCA analysis of different concentrations of MgCl2-induced phospholipids metabolomes. Each dot represents a technical replicate of samples in the plot. E. S-plot generated from OPLS-DA. Predictive component p [1] and correlation p (corr) [1] differentiated 0 and 3.125 mM MgCl2 from 50 and 200 mM MgCl2. F. Scatter plot of biomarkers in data (E). G. Diagram showing glycerophospholipid metabolism. H. qRT-PCR of the expression of genes encoding glycerophospholipid metabolism in the absence of or at the indicated concentrations of MgCl2. I. PGS and PSS levels in the absence or presence of the indicated concentrations of MgCl2. J. Activity of recombinant PSS in the absence or presence of the indicated concentrations of MgCl2. K and L Synergy analysis for MgCl2 with BLFX for ΔplsB (K) and ΔpgpA (L). Synergy is represented using a color scale or an isobologram, which compares the dose needed for 50% inhibition for synergistic agents (blue) and non-synergistic (i.e., additive) agents (red).

PE and PG are the two end products of phospholipid metabolism (Fig 5G). While expression of most genes in the phospholipid biosynthetic pathway, including psd, pldB, and glpQ, increased in the presence of 200 mM MgCl2, the expression of pssA and etuB, etuC, encoding the first enzyme and the last enzymes in the pathway, respectively, remained the same. In addition, the expression of gpsA, which encodes GpsA and transforms sn-glycerol-3P to glycerone-P, increased independent of Mg2+ concentration (Fig 5H). Phosphatidylglycerol phosphate synthase (PGS), encoded by pgsA, and phosphatidylserine synthase (PSS), encoded by pssA, are critical enzymes for the biosynthesis of PG and PE, respectively. PGS and PSS levels were elevated and reduced, respectively, with increasing Mg2+ (Fig 5I). The effect of Mg2+ on the activity of PSS was confirmed using recombinant PSS (Fig 5J). Deletion of plsB and pgpA, the first and last genes involved in PG biosynthesis, respectively, lowered cell viability in the presence of balofloxacin (note: Δpsd could not be obtained) (Fig 5K and 5L). These results indicate that phospholipids may play a role in the effect of Mg2+ on balofloxacin resistance; more specifically, PE may inhibit and PG may promote resistance to balofloxacin.

Mg2+ regulates membrane polarization, permeability and fluidity to confer balofloxacin resistance

Mg2+ remodels membrane composition, which may impair membrane potential and polarization, critical to membrane permeability and uptake of antibiotics (Lee et al., 2019; Peng et al., 2015). The voltage-sensitive dye, DiBAC4(3) showed that 12.5–200 mM MgCl2 promoted membrane depolarization in a dose-dependent manner (Fig 6A). Meanwhile, MgCl2 had a dose-dependent (Fig 6B) and time-dependent (Fig 6C) effect on proton motive force (PMF). These results suggest that Mg2+-dependent changes in membrane depolarization may influence antibiotic resistance.

Mg2+ regulates membrane polarization, permeability, and fluidity to confer balofloxacin resistance.

A and B. Depolarization (A) and PMF (B) of ATCC33787 in the absence of or at indicated concentrations of MgCl2. C and D. Dynamic depolarization (D) Membrane fluidity of ATCC33787 in the absence or presence of indicated concentrations of MgCl2, as shown by fluorescence microscopy. E. Membrane permeability of ATCC33787 in the absence or presence of the indicated concentrations of MgCl2. F and G. Membrane permeability of ATCC33787 cultured in palmitic acid (F) or linolenic acid (G) at the indicated concentrations of MgCl2. H and I. Membrane permeability of ΔfadR (H) and ΔarcA (I) in the absence or presence of the indicated concentrations of MgCl2. J. Membrane permeability of ΔpgpA in the absence or presence of the indicated concentrations of MgCl2. K. Intracellular BLFX of ATCC33787 in the presence of palmitic acid, linolenic acid, PG, and PE (left panel) or ΔfadR and ΔpgpA mutants (right panel). L. Diagram of mechanisms of Mg2+-mediated resistance to BLFX.

The efficacy of antibiotics is strongly influenced by bacterial membrane permeability and fluidity (Saeloh et al., 2018; Zhao et al., 2021). Thus, fluorescence microscopy was used to visualize these functions after incubating V. alginolyticus cells with FM5-95 and various concentrations of MgCl2. FM5-95 staining decreased with increasing concentrations of Mg2+, and no staining was observed in the presence of 200 mM Mg2+ (Fig 6D). SYTO9, a green fluorescent dye that binds to nucleic acid, enters and stains bacteria cells when there is an increase in membrane permeability (Lehtinen et al., 2004; McGoverin et al., 2020). Staining decreased with increasing MgCl2, indicating that bacterial membrane permeability declined in an Mg2+ dose-dependent manner (Fig 6E). These results indicate that exogenous Mg2+ reduces bacterial membrane permeability.

Exogenous palmitic acid also shifted the fluorescence signal peaks to the left in an MgCl2-dependent manner while palmitic acid only slightly shifted the peaks (Fig 7F). In contrast, exogenous linolenic acid shifted the peak to the right in a dose-dependent manner at 50 mM MgCl2 (Fig 6G). These results indicate that exogenous palmitic acid and linolenic acid decrease or increase membrane permeability, respectively. Membrane permeability was increased in ΔfadR cells than in WT cells and was reduced by Mg2+, remaining higher than the control (Fig 6H). However, ΔarcA cells had lower membrane permeability at lower MgCl2 concentrations (Fig 6I). ΔpgpA cells exhibited higher membrane permeability than control cells at all MgCl2 concentrations (Fig 6J). These data are consistent with the viability of the mutants described above (Figs 3L and 5L).

Relative membrane permeability appeared to correlate with relative intracellular balofloxacin and antibiotic efficacy. Exogenous palmitic acid and linolenic acid reduced and promoted the uptake of balofloxacin at 159- and 18-fold, respectively (Fig 6K). Loss of fadR and pgpA increased intracellular balofloxacin (Fig 6K). These data suggest that bacterial membrane permeability is a critical factor required for the efficacy and uptake of balofloxacin, in the presence or absence of exogenous MgCl2.

Discussion

This study explored the effect of Mg2+ on the phenotypic antibiotic resistance of V. alginolyticus. Exogenous Mg2+ was found to promote saturated fatty acid biosynthesis and inhibit unsaturated fatty acid biosynthesis. Mg2+ was also shown to upregulate PGS activity and PG abundance while downregulating PSS activity and PE abundance, decreasing membrane permeability and antibiotic uptake (Fig 6L) and enabling phenotypic resistance. These findings are consistent with the known impact of exogenous Mg2+ on bacterial survival in the presence of balofloxacin and other antibiotics, and the association between membrane permeability and drug efficacy. This study further showed that exogenous fatty acids influence membrane permeability and bacterial survival in the presence of balofloxacin.

Mg2+ is the most abundant divalent cation in cells (Pohland and Schneider, 2019; Pontes et al., 2015) playing many essential roles, including stabilizing macromolecular complexes and membranes, binding cytoplasmic nucleic acids and nucleotides, interacting with phospholipid head groups and cell surface molecules, and acting as an essential cofactor in many enzymatic reactions (Groisman et al., 2013). The present study showed for the first time that exogenous Mg2+ influences palmitic acid and linolenic acid levels and upregulates PGS while downregulating PSS. These findings extend our current understanding of the biological functions of Mg2+.

Recent findings suggest a link between fatty acid biosynthesis and the antibiotic resistance of E. tarda and V. alginolyticus (Su et al., 2021). The current study found that Mg2+ had opposing effects on the abundance of saturated and unsaturated fatty acids, stimulating saturated fatty acid biosynthesis at moderately high levels, while inhibiting unsaturated fatty acids biosynthesis at ≥3 mM. Thus, the ratio of saturated fatty acids to unsaturated fatty acids should help in predicting antibiotic resistance.

Prior studies of Gram-positive and Gram-negative bacteria (Kumariya et al., 2015; Said et al., 1987) found that 10–20 mM Mg2+ disrupts Staphylococcus aureus membranes and kills stationary-phase S. aureus cells but does not influence the survival of Escherichia coli and Bacillus subtilis (Xie and Yang, 2016). Low concentrations of Mg2+ (≤10 mM) induce PmrAB-dependent modification of lipid A in wild-type E. coli (Herrera et al., 2010). In two mundticin KS-resistant Enterococcus faecium mutants, a putative zwitterionic amino-containing phospholipid increased significantly, while phosphatidylglycerol and cardiolipin levels declined (Sakayori et al., 2003). However, the impact of Mg2+ on antibiotic resistance by regulating PGS and PSS activity was not reported. More importantly, exogenous MgCl2 was shown to target these enzymes in opposing ways to increase the PE/PG ratio, a previously unknown mechanism of Mg2+-mediated antibiotic resistance.

Prior studies show that in Stenotrophomonas maltophilia, PhoP/Q is activated by low magnesium levels and PhoP/Q inhibition or inactivation stimulates β-lactam antibiotic uptake (Huang et al., 2021). The effects of the outer membrane permeabilizers, polymyxin B nonapeptide and EDTA, are completely abolished by 3 mM Mg2+ (Kwon and Lu, 2006). In response to Mg2+-limited growth, enteric Gram-negative bacteria show higher lipid A acylation, which alters membrane permeability and reduces the uptake of cationic antimicrobial peptides (Guo et al., 1998). Mg2+ reduces the high sensitivity of rough Salmonella typhimurium, S. minnesota, and Escherichia coli 08 (which includes defects in the lipopolysaccharide carbohydrate core) mutants to several antibiotics (Stan-Lotter et al., 1979). Mg2+ (1 mM) inhibited aminoglycoside-mediated outer membrane permeabilization in Pseudomonas aeruginosa (Hancock et al., 1981). However, mechanisms underlying the impact of Mg2+ on antibiotic resistance have remained largely unknown. The current study revealed that Mg2+ plays a critical role in regulating the membrane permeability required for antibiotic resistance by modulating PE and PG during phospholipid metabolism.

Aquaculture is an environmental gateway to the development and globalization of antimicrobial resistance due to the excessive use of these drugs to prevent and treat bacterial contaminants (Cabello et al., 2020; Millanao et al., 2018). The current study found that Mg2+ increased antibiotic resistance, providing a reasonable explanation of why antibiotics are being excessively used in aquaculture and a possible solution to Mg2+-induced antibiotic resistance by balancing phospholipid metabolism. This finding should promote an overall reduction in antibacterial use that will improve environment and food safety.

In summary, Mg2+ modulated the phenotypic resistance of V. alginolyticus to balofloxacin and potentially other quinolones and other classes of antibiotics by altering membrane permeability and thus reducing antibiotic uptake. These findings inform the development of methods to improve the therapeutic effect of antibiotics on magnesium-induced phenotypic resistance.

Materials and methods

Bacterial strains and culture

V. alginolyticus ATCC33787 and V. parahaemolyticus 01 are from our laboratory collection (Jiang et al., 2022; Kou et al., 2022). They were cultured in 0.5% yeast broth (HuanKai Microbial, Guangdong, China) (pH 7.0) with 3% NaCl at 30 °C overnight. The cultures were diluted 1:50 (v/v) and grown in fresh 0.5% yeast broth supplemented with desired concentrations of MgCl2 at 30 °C until OD600 = 0.6. Bacteria were harvested by centrifugation at 8,000 rpm for 3 min and resuspended in corresponding concentrations of MgCl2 to 0.6 of OD600.

Measurement of the minimum inhibitory concentration (MIC) by microtitre-dilution-method

MIC using microtitre-dilution-method was performed as previously described (Zhang et al., 2019). In brief, the overnight bacterial cultures in 3% NaCl were diluted at 1:100 (v/v) in fresh 0.5% yeast broth supplemented with desired concentrations of MgCl2, cultured at 30 °C and collected when the cells arrived at an OD600 of 0.6 in medium without MgCl2. Then, 1 X 105 CFU cells were dispensed into each well of a 96-well microtiter polystyrene tray after which a series of 2-fold dilutions of antibiotic was added. Following incubation at 30 °C for 16 h, the MIC was defined as the lowest antibiotic concentration that inhibited visible growth. Three biological repeats were carried out.

Measurement of MIC by Oxford method

Determination of MIC by Oxford cup was performed as previously described (Liu et al., 2015). Bacterial cells cultured overnight in yeast medium were diluted at 1:100, shaken at 30 0C and 200 rpm until the OD600 nm reached at 0.6. Aliquot of 100 μL cells were spread on 0.5% yeast solid medium containing 3% NaCl and 0, 0.625, 1.25, 2.5, 5, 10, 20, or 40 mM of MgCl2. Oxford cups were placed on the solid medium, and various amounts of balofloxacin (0, 0.39, 0.78, 1.56, 3.125, 6.25, and 12.5 μg) were added. After culturing at 30 °C for 12 h, diameter of the inhibition zone was measured and the inhibition area was calculated.

Measurement of intracellular balofloxacin

Measurement of intracellular balofloxacin was performed by LC-MS analysis and plate-counting assay. For LC-MS analysis, ATCC33787 cultured in the desired concentrations of MgCl2 were collected and adjusted to OD 0.6. Aliquot of 50 mL bacterial cells were added into a 250 mL Erlenmeyer flask and then 300 μL of 10 mg/mL balofloxacin (final concentration is 60 μg/mL). After being cultured for 6 h at 30 °C, these bacterial cells were collected and adjusted to OD 1.0 with 3% NaCl containing corresponding concentrations of MgCl2. Aliquot of 30 mL bacteria were collected and washed three times using mobile phase (acetonitrile: double distilled water containing 0.1mol/l formic acid = 35:65) for LC-MS detection. The bacterial cells were added 1 mL mobile phase, crushed in ice water bath for 10 min (crush for 2S, pause for 3S at 35% power). Following by centrifugation, supernatants were collected, filtered, and then measured by using liquid chromatography for balofloxacin. Different concentrations of balofloxacin were used for a standard curve. For plate-counting assay, ATCC33787 cells cultured overnight were diluted at 1:100, shaken at 30 0C and 200 rpm until the OD600 nm reached at 0.6. The cells were diluted at 1:100 and spread on LB agar to preparing ATCC33787 plates. Meanwhile, ATCC33787 cells were cultured in medium with desired MgCl2 and at 30 °C and 200 rpm for 6 h. These cells were collected, washed and sonicated. The sonicated solution was added to Oxford cups with the ATCC33787 plates. After culturing at 30 °C for 12 h, diameter of the inhibition zone was measured. A gradient of balofloxacin was used as a standard curve for drug quantification.

Measurement of intracellular magnesium

Measurements of intracellular Mg2+ concentration were carried out as previously described with a modification (Yang et al., 2018). In brief, V. alginolyticus cultured in 0%, 0.78%, 3.125%, 12.5%, 50%, and 200% MgCl2 were collected, washed, and resuspended in the same concentrations of MgCl2 to an OD600 of 1.0. Aliquots of 20 mL of bacterial suspensions were centrifuged and washed by ddH2O once. The resulting cells were weighted, which was designated wet weight. These samples were freeze-dried overnight, which was designated dry weight. Intracellular water volume was calculated using the following formula: W × (1-W/D-0.23) as described previously (Unemoto et al., 1973). Two hundred microliters of concentrated nitric acid was added to the dried cells and then heated in 75 °C for 20 min. After diluting the samples 20-fold, they were analyzed for Mg2+ concentration by inductively coupled plasma mass spectrometry (ICP-MS) (iCAP 6500, Thermo Fisher). The intracellular Mg2+ concentration was calculated according to the volume of intracellular water.

Western blotting

Western blotting was carried out as described previously (Yao et al., 2019). Bacterial protein samples were prepared by ultrasound treatment, resolved on a 12% SDS-PAGE and transferred to nitrocellulose membranes (GE Healthcare Life Sciences). The membranes were incubated with 1:100 of the primary mouse antibodies, followed by goat anti-mouse secondary antibodies conjugated with horseradish peroxidase. Band intensities were detected by using a chemiluminescence imaging analysis system, Tanon-5200.

Metabolomics analysis

Metabolomics analysis was performed by GC-MS as described previously (Yang et al., 2018). Briefly, ATCC33787 were cultured in 0.5% yeast broth with desired MgCl2. Equivalent numbers of cells were quenched with 60% (v/v) cold methanol (Sigma) and then centrifuged at 8,000 rpm at 4 °C for 5 min. One milliliter of cold methanol was used to extract metabolites. To do this, the samples were sonicated for 5 min at a 10-Wpower setting using the Ultrasonic Processor (JY92-IIDN, Scientz, China), followed by centrifugation at 12,000 rpm in 4 °C for 10 min. Supernatants were collected and 10 µL ribitol (0.1 mg per mL, Sigma-Aldrich, USA) was added into each sample as an internal quantitative standard. The supernatants were concentrated for metabolite derivatization and then used for GC-MS analysis. Every experiment was repeated by five biological replicates. GC-MS detection and spectral processing for GC-MS were carried out using the Agilent 7890A GC equipped with an Agilent 5975C VL MSD detector (Agilent Technologies, USA) as described previously (Unemoto et al., 1973; Yang et al., 2018). Statistical difference was obtained by Kruskal Wallis test and Manne Whitney test using SPSS 13.0 and a p value < 0.01 was considered significant. Hierarchical clustering was completed in the R platform (https://cran.r-project.org/) with the function “heatmap. 2” of “gplots library”. Z score analysis was used to scale each metabolite. Multivariate statistical analysis included orthogonal partial least square discriminant analysis (OPLS-DA) implemented with SIMCA 12.0 (Umetrics, Umeå, Sweden). Control scaling was selected prior to fitting. All variables were mean centered and scaled to pareto variance of each variable. Rank-sum test and a permutation test were used to identify differentially expressed metabolites. OPLS-DA was used to reduce the high dimension of the data set. Differential metabolites to their respective biochemical pathways were outlined in the MetaboAnalyst 3.0 (http://www.metaboanalyst.ca/). Pathways were enriched by raw p value < 0.05.

qRT-PCR

Quantitative real time polymerase chain reaction (qRT-PCR) was carried out as described previously (Yang et al., 2020). Total RNA was extracted from V. alginolyticus using TRIZOL regent (Invitrogen Life Technologies) according to the protocol. RNA Electrophoresis was carried out in 1% (w/v) agarose gels to identify quality of the extracted RNA. By using a PrimeScriptTM RT reagent Kit with gDNA eraser (Takara, Japan), reverse transcription-PCR was carried out on 1 μg of total RNA and primers are listed in Table EV2. qRT-PCR was performed in 384-well plates with a total volume of 10 μL and the reaction mixtures were run on a LightCycler 480 system (Roche, Germany). Data are shown as the relative mRNA expression compared to 0% MgCl2 test with the endogenous reference 16S rRNA gene.

Antibiotic bactericidal assay

Antibiotic bactericidal assay was performed as described previously with a modification (Li et al., 2016; Zhao et al., 2021). The cultured bacteria of ATCC33787 and its mutant strains were transferred to fresh yeast medium at a dilution of 1:1000 and dispensed into test tubes and then the indicated concentrations of MgCl2 and balofloxacin were added. If desired, 2 mM 2-aminooxazole and 1 μg/mL triclosan were or a metabolite complemented. These mixtures were incubated at 30 °C and 200 rpm for 6 h. Cells were collected. To determine CFU per mL, 100-μL samples were 10-fold serially diluted with 900 μL M9 buffer and an aliquot of 5 μL of each dilution was spotted onto the LB agar plates and cultured at 30 °C for 8 h.

Construction of gene-deleted mutants

Construction of gene-deleted mutants was carried out as described previously (Kuang et al., 2021). Primers were designed as shown in Table EV3 using CE Design V1.03 software. To construct gene-deleted mutants, upstream and downstream 500-bp fragments were amplified from the genome using two pairs of primers (primers P1 and P2, primers P3 and P4), and then merged into a 1,000-bp fragment by overlap PCR using a pair of primers (primers P1 and P4). After the fragments were digested by XbaI, they were ligated into the pDS132 vector digested by the same enzymes, transformed into MC1061 competent cells. Conversion products were coated on LB plate containing 25 μg/mL chloramphenicol and cultured at 37 ℃ overnight. The plasmids from colony growing on the plate were identified by PCR using a pair of primers (primers P1 and P4) and sequenced. The sequenced plasmids were transformed into MFD λ pir competent cells as donor. MFD λpir and recipient bacterium ATCC 33787 were cultured to an optical density (OD) of 1.0 and then mixed at a ratio of 4:1. After centrifugation, the pellets were resuspended with 50 μL LB medium including DAP (100 μg/mL), dropped onto sterilized filter paper on LB medium including DAP (100 μg/mL), and cultured for 16-18 h at 37 °C.

All bacteria rinsed from the filter paper with LB medium were smeared onto the LB plate with chloramphenicol (25 μg/mL) and Ampicillin (100 μg/mL). After cultured for 16-18 h at 37 °C, bacteria were screened by LB plate with above two antibiotics. The bacteria were identified by plasmid PCR using a pair of primers P1 and P4 and sequencing and then were cultured and smeared onto the LB plates with 20% sucrose. The clones were cultured and smeared onto the LB plates with 20% sucrose again. The clones, which did not grow on the LB plates with chloramphenicol but grew on the LB plates with 20% sucrose, were identified by PCR using primer P7P8 (primer P7 is set at about 250 bp upstream of P1, and primer P8 is set at about 250 bp downstream of P4), P4P7, and P5P6 (for amplification of the target gene).

LC-MS analysis for lipidomics

ATCC33787 were cultured in medium with desired concentrations of MgCl2, washed by 3% NaCl with the desired concentrations of MgCl2, and adjusted to OD 1.0. Aliquot of 30 mL the cultures was centrifuged each sample and bacterial cells were collected. Aliquot of 800 μL distilled water was added and treated for 5 min at 100 °C to inactivate phospholipase C. Protein concentration was determined with BCA kit. Sample solution with 3 mg protein was transferred into a 10 mL centrifuge tube, and supplemented with distilled water to 800 μL. Then all samples were processed as follows: 1 mL chloroform and 2 mL methanol (to make the volume ratio of chloroform: methanol: water = 1:2: 0.8) were added and cholic acid was as an internal standard; These mixtures were vortexed for 2 min and then 1 mL chloroform was add and vortex for 30s; 1 mL 10% NaCl solution was add, vortexed for 30s, and placed at room temperature overnight. After the solution was layered, a 2.5 mL syringe was used to suck the lowest layer (chloroform layer) to a new EP tube. Solution of the chloroform layer was evaporated in the rotary evaporator for 2 h and then 1 mL mobile phase (50% and 50%B. A, 25 mM ammonium acetate / methanol (30: 70); B, methanol) was added for analysis of lipids by LC-MS. Database (https://hmdb.ca) was used for lipid identification.

Measurement of enzyme activity

Activity of pyruvate dehydrogenase (PDH), α-ketoglutarate dehydrogenase (KGDH) and succinate dehydrogenase (SDH) was carried out as previously described (Jiang et al., 2020). Cells cultured in medium were collected and washed three times with saline. The bacterial cells were suspended in Tris–HCl (pH 7.4) and disrupted by sonic oscillation for 6 min (200 W total power with 35% output, 2 s pulse, 3 s pause over ice). After centrifugation, supernatants were collected. The protein concentration of the supernatant was determined using Bradford assay (Beyotime, P0009). Then, 200 μg proteins were used for determination of pyruvate dehydrogenase (PDH), α-ketoglutarate dehydrogenase (KGDH) and succinate dehydrogenase (SDH) activity. Levels of PGS and PSS were quantified by ELISA kits according to manufacture’s instruction (Shanghai Fusheng Industrial Co., Ltd., China)

Measurement of proton motive force

Measurement of the transmembrane voltage delta PSI (δPSI), which is the electrical component of the PMF, was performed as previously described (Cheng et al., 2018). Bacteria cultured in medium with desired concentrations of MgCl2 were collected at centrifugation and labeled by DiO2(3). Approximately 1 × 107 CFU were added to a flow cytometry tube containing 1 mL buffer with 10 μM DiOC2(3) (Sigma) and incubated in the dark for 30 min at 30 0C. Samples were assayed with BD FACSCalibur flow cytometer with a 488 nm excitation wavelength. Gates for bacterial populations were based on the control population by using forward versus side scatter and red versus green emission. Size and membrane potential determined the intensity of Red (488 nm excitation, 610 nm emission) fluorescence. The diverse ratios of red and green indicated fluorescence intensity values of the gated populations. Computational formula of membrane potential: .

Measurement of depolarization

The membrane potential was estimated by measuring the fluorescence of the potential-dependent probe DiBAC4(3) (Saint-Ruf et al., 2016). DiBAC4(3) is a s voltage-sensitive probe that penetrates depolarized cells, binding intracellular proteins or membranes exhibiting enhanced fluorescence and red spectral shift. ATCC33787 were cultured in medium with desired concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacterial cells were diluted to 1 mL, and 2 μL of 5 mM DiBAC4(3) was added. After incubated for 15 min at 30 ℃ without light and vibration, these samples were filtered and detected by BD FACSCalibur.

Measurement of fluidity by fluorescence microscopy

Measurement of membrane fluidity is performed as previously described (Wen et al., 2022). Briefly, ATCC33787 were cultured in medium with indicated concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacteria cells of each sample were diluted to 1 mL and 10 μL (10 mg/mL) FM5-95 (Thermo Fisher Scientific, USA) was added. FM5-95 is a lipophilic styryl dye that insert into the outer leaflet of bacterial membrane and become fluorescence. This dye preferentially bind to the microdomains with high membrane fluidity(Wen et al., 2022). After incubated for 20 min at 30 ℃ at vibration without light, the sample was centrifuged for 10 min at 12,000 rpm. The pellets were resuspended with 20 μL of 3% NaCI. Aliquot of 2 μL sample was dropped on the agarose slide, and take photos under the inverted fluorescence microscope.

Measurement of membrane permeability

Measurement of membrane permeability was carried out as described previously (Su et al., 2021). ATCC33787 were cultured in medium with desired concentrations of MgCl2, collected and then adjusted to OD 0.6. Aliquot of 100 μL bacteria cells of each sample were diluted to 1 mL and 2 μL 10 mg/mL SYT09 was added. After incubated for 15 min at 30 ℃ at vibration without light, the mixtures were filtered and measured by flow cytometry (BD FACSCalibur, USA).